---
title: "ST558_HW4_Kayla"
author: "Kayla Nowak"
format: html
editor: visual
---

# Task 1

## 1.

Rewrite the following nested function using baseR's chaining operator:

```{r}
library(tidyverse)

#Original Code:
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

#Rewritten with baseR (as data.frame):
colnames(iris)

iris[order(iris$Species), 3:5] |>   #arrange, select   
  subset(Petal.Length < 1.55)       #filter (but row names do not change)

```

## 2.

**What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?**

Long data has 1 row per observation and 1 col per variable vs. wide data has multiple observations stored as different columns for a single row.

Long format data is preferred for statistical analysis.

# Task 2

## 3-4.

Revisiting the Glass data from HW 3:

```{r}
glass <- read_delim("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", 
                                  "K", "Ca", "Ba", "Fe", "Type_of_glass"))
```

Overwrite `Type_of_glass` as a factor using mutate().

Then continue the chain to keep only observations where `Fe` \< 0.2 and the type of glass is "tableware" or "headlamps".

```{r}
glass |>
  mutate(Type_of_glass = factor(Type_of_glass, levels=c(1:7), 
                                labels=c("building_windows_float_processed",
                                         "building_windows_non_float_processed",
                                         "vehicle_windows_float_processed",
                                         "vehicle_windows_non_float_processed",
                                         "containers",
                                         "tableware",
                                         "headlamps"))) |>
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps")) |> 
  print(n=100)
```

## 5-6.

Revisiting the Yeast data from HW 3:

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                                  "erl", "pox", "vac", "nuc", "class"))
```

Remove `seq_name` and `nuc` columns.

Then continue the chain to add columns corresponding to the mean and median of each numeric variable at each class grouping.

***!! Come back to this later... across() deprecation warning !!***

```{r}
#yeast |>
  #select(-c(seq_name, nuc)) |>
  #group_by(class) |>
  #mutate(across(where(is.numeric), mean, names= "{.col}_mean"))
                #list(mean=mean, median=median), 
                #names= "{.col}_{.fn}") )
```

## 7-10.

Revisiting the combined Wine data from HW 3:

```{r}
library("readxl")
library("readr")

#read variable names
wine_vars  <- read_xlsx(path="../HW3/white-wine.xlsx", sheet = "variables") 

#white wine: read and format
white_wine <- read_xlsx(path="../HW3/white-wine.xlsx", sheet="white-wine") |>
  setNames(wine_vars$Variables) |>
  mutate(type="white")

#red wine: read and format
red_wine  <- read_delim(file="../HW3/red-wine.csv", delim=";") |>
  setNames(wine_vars$Variables) |>
  mutate(type="red")

#stack data
all_wine <- bind_rows(white_wine, red_wine)
all_wine
```


Filter to high quality wines (`quality` > 6.5) with reasonable alcohol value (`alcohol` < 132).

Then sort the data from highest quality to lowest.

Then select the variables containing `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.

Then add mean and standard deviation of `alcohol` at each setting of `quality` to the dataset.

```{r}
all_wine |>
  filter(quality > 6.5 & alcohol < 132) |>
  arrange(desc(quality)) |>   #descending for highest to lowest
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(alcohol_mean = mean(alcohol),
         alcohol_sd = sd(alcohol))
```