---
title: "ST558_HW4_Kayla"
author: "Kayla Nowak"
format: html
editor: visual
---

# Task 1

## 1.

Rewrite the following nested function using baseR's chaining operator:

```{r}
library(tidyverse)

#Original Code:
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

#Rewritten with baseR (as data.frame):
colnames(iris)

iris[order(iris$Species), 3:5] |>   #arrange, select   
  subset(Petal.Length < 1.55)       #filter (but row names do not change)

```

## 2.

**What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?**

Long data has 1 row per observation and 1 col per variable vs. wide data has multiple observations stored as different columns for a single row.

Long format data is preferred for statistical analysis.

# Task 2

## 3-4.

Revisiting the Glass data from HW 3:

```{r}
glass <- read_delim("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", 
                                  "K", "Ca", "Ba", "Fe", "Type_of_glass"))
```

Overwrite `Type_of_glass` as a factor using mutate().

Then continue the chain to keep only observations where `Fe` \< 0.2 and the type of glass is "tableware" or "headlamps".

```{r}
glass |>
  mutate(Type_of_glass = factor(Type_of_glass, levels=c(1:7), 
                                labels=c("building_windows_float_processed",
                                         "building_windows_non_float_processed",
                                         "vehicle_windows_float_processed",
                                         "vehicle_windows_non_float_processed",
                                         "containers",
                                         "tableware",
                                         "headlamps"))) |>
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps")) |> 
  print(n=100)
```

## 5-6.

Revisiting the Yeast data from HW 3:

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                                  "erl", "pox", "vac", "nuc", "class"))
```

Remove `seq_name` and `nuc` columns.

Then continue the chain to add columns corresponding to the mean and median of each numeric variable at each class grouping.

*Note: switched to separate mean() and median() calls due to the across() deprecation warning. Using the `\(x) fn(x, na.rm=TRUE)` alternative code from the warning message.*

```{r}
yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(where(is.numeric), 
                .fns = list(mean = \(x) mean(x, na.rm=TRUE), 
                            median = \(x) median(x, na.rm=TRUE)), 
                .names = "{.col}_{.fn}") ) #|> view()
```

## 7-10.

Revisiting the combined Wine data from HW 3:

```{r}
library("readxl")
library("readr")

#read variable names
wine_vars  <- read_xlsx(path="../HW3/white-wine.xlsx", sheet = "variables") 

#white wine: read and format
white_wine <- read_xlsx(path="../HW3/white-wine.xlsx", sheet="white-wine") |>
  setNames(wine_vars$Variables) |>
  mutate(type="white")

#red wine: read and format
red_wine  <- read_delim(file="../HW3/red-wine.csv", delim=";") |>
  setNames(wine_vars$Variables) |>
  mutate(type="red")

#stack data
all_wine <- bind_rows(white_wine, red_wine)
all_wine
```

Filter to high quality wines (`quality` \> 6.5) with reasonable alcohol value (`alcohol` \< 132).

Then sort the data from highest quality to lowest.

Then select the variables containing `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.

Then add mean and standard deviation of `alcohol` at each setting of `quality` to the dataset.

```{r}
all_wine |>
  filter(quality > 6.5 & alcohol < 132) |>
  arrange(desc(quality)) |>   #descending for highest to lowest
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(alcohol_mean = mean(alcohol),
         alcohol_sd = sd(alcohol))
```

# Task 3.

Working with the `Lahman.db`

## 11.

Connect to the database and look at all the tables in the database.

```{r}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "lahman.db")

#list all the tables
dbListTables(con)
```

## 12.

Return all of the data from the `Teams` table for the year 2015 using `tbl()` and `dplyr`.

```{r}
tbl(con, "Teams") |> 
  filter(yearID == 2015) |>
  collect() #force full return
```

## 13.

Repeat this process using `sql()` code within `tbl()`.

```{r}
tbl(con, sql(
"SELECT *
FROM `Teams`
WHERE (`yearID` = 2015.0)")
) |>
  collect()
```

## 14.

Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category.

```{r}
tbl(con, "HallOfFame") |>
  filter(inducted=="Y") |>                  #only those inducted
  select(playerID, yearID, category) |>     #fields to keep
  collect()
```

## 15.

Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added to what was returned above.

```{r}
left_join(tbl(con, "HallOfFame") |> filter(inducted=="Y"),
          tbl(con,"People"), 
          by= c("playerID")) |>
  select(nameFirst, nameLast, playerID, yearID, category) |>
  collect()
```

## 16.

Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table.

Then determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers.

Create a new variable that is the career win/loss percentage.

Sort the resulting data by win/loss percentage variable (from largest to smallest).

```{r}
tbl(con, "Managers") |>
  select(playerID, G, W, L) |> collect() |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE),
            Career_WL_Pct = Total_W/G_managed) |>
  arrange(desc(Career_WL_Pct))
```

## 17.

Return people who managed a team and were inducted into the hall of fame (regardless of their hall of fame category). Return all 9 columns used in the previous 2 questions.

*Note: Summarize function drops anything not in the group_by() statement. As such, tackle the Manager data first, then join with People and Hall of Fame data. Summarize also requires collect() to be called prior to to summarizing. This means that each data element for the join must \|\> collect() to ensure X:Y source agreement.*

*Note: This code completes the full return in one big, nested join. Alternatively, each database return could be saved as its own R object (tibble), and then dplyr join could simply reference the objects.*

```{r}

inner_join(
  
  #Top X = Managers + names from People
  left_join(#Nested X = Managers, fully processed 
            tbl(con, "Managers") |>
              select(playerID, G, W, L) |> collect() |>
              group_by(playerID) |>
              summarize(G_managed = sum(G, na.rm = TRUE),
                        Total_W = sum(W, na.rm = TRUE),
                        Total_L = sum(L, na.rm = TRUE),
                        Career_WL_Pct = Total_W/G_managed) |>
              arrange(desc(Career_WL_Pct)),
            #Nested Y = People, just to pick up names
            tbl(con,"People") |> 
              select(playerID, nameFirst, nameLast) |> 
              collect(),
            #Nested by playerID (for left join)
            by="playerID"),
  
  #Top Y = Hall of Fame, fully processed
  tbl(con, "HallOfFame") |> 
    filter(inducted=="Y") |> 
    select(playerID, yearID, category) |>
    collect(),
  
  #Top By playerID (for inner join)
  by="playerID") |>
  #Order Variables
  select(nameFirst, nameLast, playerID, yearID, category, everything())


## Example alternate code (if objects had been created):
#inner_join(inducted_people, managers, by="playerID")

```

Disconnect from database.

```{r}
dbDisconnect(con)
```
