---
title: "ST558 Homework 4"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  Rewriting a nested function call using `baseR` chaining

Without chaining, the function looks like this (Have to read from inside out, confusing!)

```{r}
#| message: False
#| warning: False
library(tidyverse)
```

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)
```

Using chaining, we can re-write the function like so:

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

This gives us the same result, and is much easier to follow!

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

    Long format data is structured with more rows, where each row consists of one observation and each column consists of one variable. Wide format data has a single row per subject and may have more than one observation in a given row (more columns). Generally we prefer long format as most software is built to deal with data formatted this way.

# Task 2: Manipulating data

## Glass Data

We'll use the `glass` data from homework 3 again, so let's re-read in that data.

Dataset available here: <https://www4.stat.ncsu.edu/~online/datasets/glass.data.>

```{r}
glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                   "Ca", "Ba", "Fe", "Type"))
# Print tibble
glass
```

3.  Start a chain on your data object the overwrites the `Type` variable using `mutate()`

    We'll start by creating a `factor` version the variable that replaces the numbers with string labels.

4. Continue the chain and keep only observations where the `Fe` variable is less than 0.2 and the Type of glass
is either tableware or headlamp
```{r}
glass |>
  mutate(Type = factor(Type, levels = c(1:7), # 3. Converting type to factor strings
                       labels=c("building_windows_float_processed",
                                "building_windows_non_float_processed",
                                "vehicle_windows_float_processed",
                                "vehicle_windows_non_float_processed",
                                "containers",
                                "tableware",
                                "headlamps"))) |> 
  filter(Fe < 0.2, Type %in% c("tableware", "headlamps")) # 4. Filter for select observations
```

Now we can see the type is displayed as our factor strings ("tableware" instead of 1), and
only 38 rows are returned after applying these filters (only the first 10 are displayed).

## Yeast Data

We'll use the Yeast data, also from homework 3 and real it into R directly from the URL here: <https://www4.stat.ncsu.edu/~online/datasets/yeast.data.>

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
    col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac",
             "nuc", "class"))

# Print tibble
yeast
```

5. Start a chain that removes the `seq_name` and `nuc` columns

6. Continue your chain to add columns corresponding to the mean and median of each numeric variable
(mcg, gvh, alm, mit, erl, pox, and vac) at each `class` grouping using the `across()` function
```{r}
yeast |>
  select(-c(seq_name, nuc)) |> # 5. removing columns
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean=mean, median=median), # 6. Adding mean and median columns
                .names= "{.col}_{.fn}"))
```

## Wine Data

In Homework 3 we read in two wine data sets and combine them, here we'll repeat that process.

```{r}
#| message: False
#| warning: False

library(readxl)
```

```{r}
# Read variable names
wine_vars  <- read_xlsx(path="../HW3/white-wine.xlsx", sheet = "variables") 

# White wine
white_wine <- read_xlsx(path="../HW3/white-wine.xlsx", sheet="white-wine") |>
  setNames(wine_vars$Variables) |>
  mutate(type="white")

# Red wine
red_wine  <- read_delim(file="../HW3/red-wine.csv", delim=";") |>
  setNames(wine_vars$Variables) |>
  mutate(type="red")

# Combine data
total_wine <- bind_rows(white_wine, red_wine)
total_wine
```

7. Start a chain on the new combined data object to filter the data to only look at
high quality wines (quality > 6.5) and wines with a reasonable alcohol value (alcohol < 13.2)

8. Continue the chain to sort the data from highest `quality` to lowest

9. Continue the chain to select only the variables that contain `acid`, the `alcohol` variable, the `type`
variable, and the `quality` variable

10. Finally, add the mean and standard deviation of the `alchohol` variable to the data set for each setting of the `quality` variable

```{r}
total_wine |>
  filter(quality > 6.5, alcohol < 13.2) |> # Step 7
  arrange(desc(quality)) |> # Step 8
  select(contains("acid"), alcohol, type, quality) |> # Step 9
  group_by(quality) |> # Step 10
    mutate(alcohol_mean = mean(alcohol),
         alcohol_sd = sd(alcohol))
```

## Task 3: Database practice

First, let's download the `Lahman.db` file associated with the connecting to databases notes

11. Connect to the database and then look at all of the tables in the database

```{r}
#| message: False
#| warning: False

library(RSQLite)
library(DBI)
```


```{r}
con <- dbConnect(RSQLite::SQLite(), "https://www4.stat.ncsu.edu/~online/datasets/lahman.db")
dbListTables(con)

# Need to figure out how to connect to the database, can't add the .db file to github HW4 folder as its too large (31 MB). Do I just need to add It to the Rproject folder locally?
```

