---
title: "ST558 Homework 4"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  Rewriting a nested function call using `baseR` chaining

Without chaining, the function looks like this (Have to read from inside out, confusing!)

```{r}
#| message: False
#| warning: False
library(tidyverse)
```

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)
```

Using chaining, we can re-write the function like so:

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

This gives us the same result, and is much easier to follow!

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

    Long format data is structured with more rows, where each row consists of one observation and each column consists of one variable. Wide format data has a single row per subject and may have more than one observation in a given row (more columns). Generally we prefer long format as most software is built to deal with data formatted this way.

# Task 2: Manipulating data

## Glass Data

We'll use the `glass` data from homework 3 again, so let's re-read in that data.

Dataset available here: <https://www4.stat.ncsu.edu/~online/datasets/glass.data.>

```{r}
glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                   "Ca", "Ba", "Fe", "Type"))
# Print tibble
glass
```

3.  Start a chain on your data object the overwrites the `Type` variable using `mutate()`

    We'll start by creating a `factor` version the variable that replaces the numbers with string labels.

4. Continue the chain and keep only observations where the `Fe` variable is less than 0.2 and the Type of glass
is either tableware or headlamp
```{r}
glass |>
  mutate(Type = factor(Type, levels = c(1:7), # 3. Converting type to factor strings
                       labels=c("building_windows_float_processed",
                                "building_windows_non_float_processed",
                                "vehicle_windows_float_processed",
                                "vehicle_windows_non_float_processed",
                                "containers",
                                "tableware",
                                "headlamps"))) |> 
  filter(Fe < 0.2, Type %in% c("tableware", "headlamps")) # 4. Filter for select observations
```

Now we can see the type is displayed as our factor strings ("tableware" instead of 1), and
only 38 rows are returned after applying these filters (only the first 10 are displayed).

## Yeast Data

We'll use the Yeast data, also from homework 3 and real it into R directly from the URL here: <https://www4.stat.ncsu.edu/~online/datasets/yeast.data.>

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
    col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac",
             "nuc", "class"))

# Print tibble
yeast
```

5. Start a chain that removes the `seq_name` and `nuc` columns

6. Continue your chain to add columns corresponding to the mean and median of each numeric variable
(mcg, gvh, alm, mit, erl, pox, and vac) at each `class` grouping using the `across()` function
```{r}
yeast |>
  select(-c(seq_name, nuc)) |> # 5. removing columns
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean=mean, median=median), # 6. Adding mean and median columns
                .names= "{.col}_{.fn}"))
```

## Wine Data