---
title: "ST558 Homework 4 - Group 1"
author: 
  - name: Kayla Nowak
  - name: Ryan Merserau
  - name: Will Carter
  - name: Yefrid Cordoba
format: html
---

## Task 1:Conceptual Questions

### 1.

Rewriting a nested function call using `baseR` chaining

Without chaining, the function looks like this (Have to read from inside out, confusing!)

```{r}
#| message: False
#| warning: False
library(tidyverse)
library(readxl)
library(DBI)
library(RSQLite)
library(readr)
```

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)
```

Using chaining, we can re-write the function like so:

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

This gives us the same result, and is much easier to follow!

### 2.

Wide format gives each variable its own column with one row per unique record. Long format gives each row a single observation so one subject spans multiple rows instead of multiple columns. Long Format is generally preferred for analysis. Modeling functions expect one row per observation so long format is more intuitive for this.

## Task 2: Chaining Practice

### Glass data

We'll use the `glass` data from homework 3 again, so let's re-read in that data.

Dataset available here: <https://www4.stat.ncsu.edu/~online/datasets/glass.data.>

```{r}
glass <- read_delim("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", 
                                  "K", "Ca", "Ba", "Fe", "Type_of_glass"))
```

#### 3 & 4

Overwrite `Type_of_glass` as a factor using mutate().

Then continue the chain to keep only observations where `Fe` \< 0.2 and the type of glass is "tableware" or "headlamps".

```{r}
glass |>
  mutate(Type_of_glass = factor(Type_of_glass, levels = c(1:7), # 3. Converting type to factor strings
                       labels=c("building_windows_float_processed",
                                "building_windows_non_float_processed",
                                "vehicle_windows_float_processed",
                                "vehicle_windows_non_float_processed",
                                "containers",
                                "tableware",
                                "headlamps"))) |> 
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps")) # 4. Filter for select observations
```

Now we can see the type is displayed as our factor strings ("tableware" instead of 1), and only 38 rows are returned after applying these filters (only the first 10 are displayed).

### Yeast data

We'll use the Yeast data, also from homework 3 and real it into R directly from the URL here: <https://www4.stat.ncsu.edu/~online/datasets/yeast.data.>

```{r}
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                       col_names = c("seq_name", "mcg", "gvh", "alm", "mit", 
                                     "erl", "pox", "vac", "nuc", "class"))
```

#### 5 & 6

Remove `seq_name` and `nuc` columns.

Then continue the chain to add columns corresponding to the mean and median of each numeric variable at each class grouping.

```{r}
yeast_data |>
  group_by(class) |>
  mutate(seq_name = NULL,
         nuc = NULL,
         across(where(is.numeric), 
                list(mean = mean, median = median),
                .names = "{.col}_{.fn}"))
```

### Wine data

In Homework 3 we read in two wine data sets and combine them, here we’ll repeat that process.

```{r}
# Read variable names
wine_vars  <- read_xlsx(path="../HW3/white-wine.xlsx", sheet = "variables") 

# White wine
white_wine <- read_xlsx(path="../HW3/white-wine.xlsx", sheet="white-wine") |>
  setNames(wine_vars$Variables) |>
  mutate(type="white")

# Red wine
red_wine  <- read_delim(file="../HW3/red-wine.csv", delim=";") |>
  setNames(wine_vars$Variables) |>
  mutate(type="red")

# Combine data
wine_data <- bind_rows(white_wine, red_wine)
wine_data
```

#### 7 - 10

Filter to high quality wines (`quality` \> 6.5) with reasonable alcohol value (`alcohol` \< 132).

Then sort the data from highest quality to lowest.

Then select the variables containing `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.

Then add mean and standard deviation of `alcohol` at each setting of `quality` to the dataset.

```{r}
wine_data |>
  filter(quality > 6.5 & alcohol < 132) |>
  arrange(desc(quality)) |>   #descending for highest to lowest
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(alcohol_mean = mean(alcohol),
         alcohol_sd = sd(alcohol))
```

## Task 3: Database Practice

First, let's download the `Lahman.db` file associated with the connecting to databases notes.

### 11.

We connect to the database and then look at all of the tables in the database.

```{r}
con <- dbConnect(RSQLite::SQLite(), "lahman.db")

#list all the tables
dbListTables(con)
```

### 12.

Return all of the data from the `Teams` table for the year 2015 using `tbl()` and `dplyr`.

```{r}
tbl(con, "Teams") |>
  filter(yearID == 2015)
```

### 13. 

Repeat this process using `sql()` code within `tbl()`.

```{r}
tbl(con, sql(
  "SELECT *
  FROM `Teams`
  WHERE (`yearID` = 2015.0)")
)
```

### 14.

Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category.

```{r}
tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)
```

### 15.

Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added

```{r}
hof_names <- tbl(con, "HallOfFame") |> 
                filter(inducted == "Y") |> 
                select(playerID, yearID, category) |> 
                left_join(tbl(con, "People") |> 
                select(playerID, nameFirst, nameLast), by = "playerID") |>
                collect()
hof_names
```

### 16.

Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table.

Then determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers.

Create a new variable that is the career win/loss percentage.

Sort the resulting data by win/loss percentage variable (from largest to smallest).

```{r}
manager_records <- tbl(con, "Managers") |> 
                      select(playerID, G, W, L) |>
                      group_by(playerID) |>
                      summarize(G_managed = sum(G, na.rm = TRUE),
                                Total_W = sum(W, na.rm = TRUE),
                                Total_L = sum(L, na.rm = TRUE)) |>
                      collect() |>
                      mutate(win_loss = Total_W / G_managed) |>
                      arrange(desc(win_loss))
manager_records
```

### 17.

Return people who managed a team and were inducted into the hall of fame (regardless of their hall of fame category). Return all 9 columns used in the previous 2 questions.

```{r}
hof_managers <- hof_names |>
  inner_join(manager_records, by = "playerID")

hof_managers
```

Disconnect from database.

```{r}
dbDisconnect(con)
```