---
title: "ST558_HW4_Yefrid"
author: "Yefrid Cordoba"
format: html
---

## Task 1: Conceptual questions

### 1.

Writing the following nested function using the `baseR` chaining operator.
```{r}
#| eval: False

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)

```

Now using the pipe operator

```{r}
#| eval: False

as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

### 2.

The long format data is the one that will increase the number of rows and decrease the number or columns, while the wide format data, widens the data by increasing the number of columns and decreasing the number of rows.    

For statistical analysis is preferred to use the long format data, since  we can get in each row an observation.

## Task 2:

### Glass data

We copy the code from HW3
```{r}
#| message: False
#| 
library(tidyverse)
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                     "Ca", "Ba", "Fe", "Type"))
```

#### 3 & 4.

Now we are going to overwrite the column `Type` as `Type_of_glass`, and replace the numbers to the corresponding type as factor.

```{r}
glass_data |>
  mutate(Type_of_glass = factor(Type,
                                labels = c("building_windows_float_processed",
                                         "building_windows_non_float_processed",
                                         "vehicle_windows_float_processed",
                                         "containers",
                                         "tableware",
                                         "headlamps")),
         Type = NULL) |>
  filter(Fe < 0.2,
         Type_of_glass %in% c("tableware", "headlamps")) |>
  show()
```

### Yeast data

We get the code from HW3

```{r}
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                       col_names = c("seq_name", "mcg", "gvh", "alm", "mit", 
                                     "erl", "pox", "vac", "nuc", "class"))
```

#### 5 & 6.

We are going to remove two columns using the mutate function, and add the columns to the mean and median for each numeric variable.

```{r}
yeast_data |>
  group_by(class) |>
  mutate(seq_name = NULL,
         nuc = NULL,
         across(where(is.numeric), 
                list(mean = mean, median = median),
                .names = "{.col}_{.fn}")) |>
  show()
  
```

### Wine data

From HW3

```{r}
#| message: False
#| warning: False

library(readxl)

white_data <- read_excel("../HW3/white-wine.xlsx")

var_wine <- read_excel("../HW3/white-wine.xlsx",
                       sheet = excel_sheets("../HW3/white-wine.xlsx")[2])
#And we re-assign the colnames to the tibble with the data

colnames(white_data) <- var_wine$Variables #it is necessary to get the variable column as a vector of characters

white_data$type <- "white"

red_data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";",
                       col_names = var_wine$Variables,
                       skip = 1
                      )
red_data$type <- "red"

wine_data <- dplyr::bind_rows(white_data, red_data)
```

#### 7, 8, 9, & 10.

We start by taking the wine data already created, and 

```{r}
wine_data |>
  filter(quality > 6.5,
         alcohol < 13.2) |> # We get the vales that have a high quality and reasonable alcohol content
  arrange(desc(quality)) |> #then we sorted in descending order on the quality variable
  select(contains("acid"), alcohol, quality, type) |> #just select the columns with acid, alcohol, quality, and type
  group_by(quality) |> #we group by quality to get means per quality group.
  mutate(alcohol_mean = mean(alcohol)) |>
  show()
```

## Task 3: Database practice
