---
title: "ST558_HW2_Yefrid"
author: "Yefrid Cordoba"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1

First we create two vectors for the values of the blood pressure pre and post-treatment measurements for 20 patients
```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2

Now we assign names to the elements to easily match both subjects before and after treatment.
```{r}
ns <- paste("Subject", 1:20, sep = "_") # we create the names in order for each of the subjects

#And we assign them to each of the names attributes for the two vectors

names(pre) <- ns

names(post) <- ns
```

### 3

Here we need to calculate the change in blood pressure for each patient.
```{r}
ch <- pre - post
```

### 4

And we calculate the average decrease in blood pressure across all patients.
```{r}
mean(ch)
```

### 5

Using the `which()` function, we extract the indices and names of the subjects when the blood pressure decreased after treatment.
```{r}
which(ch > 0)
```

### 6

To subset the vector with only the ones that had a positive change, we use the function `which()` to get the indices when the condition is met.
```{r}
dec <- ch[which(ch > 0)]

dec
```

### 7

And finally we calculate the average decrease in blood pressure for just the patients that actually had a decrease in blood pressure.
```{r}
mean(dec)
```

## Task 2: Basic Data Frame Practice

### 1

We are going to create a data frame with the corresponding data for patient, pre, post, and difference in blood pressure.
```{r}
bp_data <- data.frame(ns, pre, post, ch)

names(bp_data) <- c("patient", "pre_bp", "post_bp", "diff_bp")

bp_data
```

### 2

Now we just return the rows when the difference of blood pressure is positive (decrease in blood pressure).
```{r}
subset(bp_data, diff_bp > 0) #Returns a new data frame that meet the conditions
```

### 3

Now we add a column to the data frame of boolean when the post-treatment blood pressure is less than 120.
```{r}
bp_data$low_bp <- bp_data$post_bp < 120
```

### 4 

Finally, we print the data frame nicely.

```{r}
knitr::kable(bp_data)
```

## Task 3: List Practice

