---
title: "ST558 HW2 (Ryan)"
author: "Ryan Mersereau"
format: html
editor: visual
---

# Task 1: Basic Vector practice

First, we create two vectors corresponding to the blood pressure of 20 subjects. The first, `pre_bp` is pre-treatment measurements, and the second, `post_bp` is post-treatment measurements.

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 
            128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119,
             130, 122, 106, 106, 124, 102, 117, 113)

is.vector(pre_bp)
```

Next, we will assign `names` to the vector elements using this `paste()` function

```{r}
patient <- c(paste("Subject", 1:20, sep = "_"))
names(pre_bp) <- patient
names(post_bp) <- patient
print(head(pre_bp))
```

We can calculate the change in blood pressure by subtracting post-treatment measurements from pre-treatment measurements.

```{r}
diff_bp <- pre_bp - post_bp
names(diff_bp) <- patient
```

Now, we can calculate the average decrease in blood pressure across all patients

```{r}
mean(diff_bp)
```

From this data, blood pressure decreased 17 bpm, on average, after treatment.

We can specifically determine which patients experienced a decrease in blood pressure after treatment (positive change) using the `which()` function

```{r}
which((diff_bp) > 0)
```

We can see that all patients experienced a decrease in blood pressure after treatment except for subject 9, 13, and 17. Therefore, we can make a vector that is a subset of the differences to only include subjects whose blood pressure decreased.

```{r}
positive_diff_bp <- diff_bp[which((diff_bp) > 0)]
print(positive_diff_bp)
```

Finally, we can recalculate the average decrease in blood pressure from this subset

```{r}
mean(positive_diff_bp)
```

From this subset of only subjects who experienced a decrease in blood pressure after treatment, the average decrease for this sample was 20.647 bpm

# Task 2: Basic Data Frame practice

Continuing from task 1, we'll create a data frame object with four columns corresponding to: `patient`, `pre_bp`, `post_bp`, and `diff_bp`

```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df
```

Next, let's return only rows where `diff_bp` is negative, we'll use the`[` function

Using `bp_df[ , ]`, we can subset rows (first argument) or columns. By specifying `bp_df[4] < 0` in the rows argument, we are subsetting for rows where the fourth column of our dataframe, `diff_bp` is less than 0.

```{r}
bp_df[bp_df[4] < 0, ]
```

Next, let's add a new boolean column to our dataframe that returns TRUE if `post_bp` is less than 120. We can do this by using `$` to reference a column.

```{r}
bp_df$low_post <- bp_df$post_bp < 120
head(bp_df)
```

Here we referenced the `post_bp`, column, and by giving it the argument of \< 120, this returns TRUE if the conditions are met for this column, and FALSE if not. We then can assign these boolean values to a new column in our dataframe, which I've named `low_post`

Finally, let's print the data frame out nicely in our final document

```{r}
knitr::kable(bp_df, caption = "Blood Pressure by Subject")
```

# Task 3: List practice

Now, we have similar data from another experiment where the treatment was actually a placebo.

We'll have new data for pre and post treatment values for 10 subjects, which we'll store in a dataframe structured similarly to the dataframe from task 2

```{r}
placebo_pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_diff_bp <- placebo_pre_bp - placebo_post_bp
placebo_patient <- c(paste("Subject", 1:10, sep = "_"))
```

```{r}
placebo_bp_df <- data.frame(placebo_patient, placebo_pre_bp, placebo_post_bp, placebo_diff_bp)
placebo_bp_df$low_post <- placebo_bp_df$placebo_post_bp < 120
placebo_bp_df
```

Now, we can create and store a list with two elements:

- 1st element: `treatment`, our data frame from task 2
- 2nd element: `placebo`, our data frame created above

```{r}
my_list <- list(treatment = bp_df, placebo = placebo_bp_df)
my_list
```

We can access the first list element in three different ways:
