---
title: "ST558 HW1 (Ryan)"
author: "Ryan Mersereau"
format: html
editor: visual
---

# Task 1: Basic Vector practice

First, we create two vectors corresponding to the blood pressure of 20 subjects. The first, `pre_bp` is pre-treatment measurements, and the second, `post_bp` is post-treatment measurements.

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 
            128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119,
             130, 122, 106, 106, 124, 102, 117, 113)

is.vector(pre_bp)
```

Next, we will assign `names` to the vector elements using this `paste()` function

```{r}
names(pre_bp) <- c(paste("Subject", 1:20, sep = "_"))
names(post_bp) <- c(paste("Subject", 1:20, sep = "_"))
print(head(pre_bp))
```

We can calculate the change in blood pressure by subtracting post-treatment measurements from pre-treatment measurements.

```{r}
diff_bp <- pre_bp - post_bp
names(diff_bp) <- c(paste("Subject", 1:20, sep = "_"))
```

Now, we can calculate the average decrease in blood pressure across all patients

```{r}
mean(diff_bp)
```

From this data, blood pressure decreased 17 bpm, on average, after treatment.

We can specifically determine which patients experienced a decrease in blood pressure after treatment (positive change) using the `which()` function

```{r}
which((diff_bp) > 0)
```

We can see that all patients experienced a decrease in blood pressure after treatment except for subject 9, 13, and 17. Therefore, we can make a vector that is a subset of the differences to only include subjects whose blood pressure decreased.

```{r}
positive_diff_bp <- diff_bp[which((diff_bp) > 0)]
print(positive_diff_bp)
```

Finally, we can recalculate the average decrease in blood pressure from this subset

```{r}
mean(positive_diff_bp)
```

From this subset of only subjects who experienced a decrease in blood pressure after treatment, the average decrease for this sample was 20.647 bpm
