---
title: "ST558_HW2_Yefrid"
author: "Yefrid Cordoba"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1

First we create two vectors for the values of the blood pressure pre and post-treatment measurements for 20 patients
```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2

Now we assign names to the elements to easily match both subjects before and after treatment.
```{r}
ns <- paste("Subject", 1:20, sep = "_") # we create the names in order for each of the subjects

#And we assign them to each of the names attributes for the two vectors

names(pre) <- ns

names(post) <- ns
```

### 3

Here we need to calculate the change in blood pressure for each patient.
```{r}
ch <- pre - post
```

### 4

And we calculate the average decrease in blood pressure across all patients.
```{r}
mean(ch)
```

### 5

Using the `which()` function, we extract the indices and names of the subjects when the blood pressure decreased after treatment.
```{r}
which(ch > 0)
```

### 6

To subset the vector with only the ones that had a positive change, we use the function `which()` to get the indices when the condition is met.
```{r}
dec <- ch[which(ch > 0)]

dec
```

### 7

And finally we calculate the average decrease in blood pressure for just the patients that actually had a decrease in blood pressure.
```{r}
mean(dec)
```

## Task 2: Basic Data Frame Practice

### 1

We are going to create a data frame with the corresponding data for patient, pre, post, and difference in blood pressure.
```{r}
bp_data <- data.frame(ns, pre, post, ch, row.names = NULL)

names(bp_data) <- c("patient", "pre_bp", "post_bp", "diff_bp")

bp_data
```

### 2

Now we just return the rows when the difference of blood pressure is positive (decrease in blood pressure).
```{r}
subset(bp_data, diff_bp > 0) #Returns a new data frame that meet the conditions
```

### 3

Now we add a column to the data frame of boolean when the post-treatment blood pressure is less than 120.
```{r}
bp_data$low_bp <- bp_data$post_bp < 120
```

### 4 

Finally, we print the data frame nicely.

```{r}
knitr::kable(bp_data)
```

## Task 3: List Practice

### 1

We create a data frame as the previous one with the new subjects that received the placebo treatment.
```{r}
#Create the vectosr with the data
pre_p <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

post_p <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

#Create the Subjects names
Sub_p <- paste("Subject", 1:10, sep = "_")

#Create the difference column
diff_p <- pre_p - post_p

#Create the vector of boolean when post placebo is lower than 120
low_p <- post_p < 120
```

And now we put everything in a data frame and print it out.
```{r}
bp_pla <- data.frame(Sub_p, pre_p, post_p, diff_p, low_p)

names(bp_pla) <- c("patient", "pre_bp", "post_bp", "diff_bp", "low_bp")

knitr::kable(bp_pla)
```

### 2

We create a list with the two data frames previously created, and assign name to each one.

```{r}
bp_study <- list(treatment = bp_data, placebo = bp_pla)
```

### 3

After creating the list, we will access the first element in three different ways.

```{r}
# Using the indices in the list
bp_study[1] 

#Using the names assigned and brackets
bp_study['treatment']

#Using the operator $
bp_study$treatment
```

### 4

And now in one line of code we will access the `pre_bp` column from the `placebo` data frame.

```{r}
bp_study[["placebo"]]["pre_bp"]
```

## Task 4: Control Flow Practice

### 1