---
title: "Homework 2 - ST558 Group 1"
author: "Ryan Mersereau, Kayla Nowak, Yefrid Cordoba Ortiz, Will Carter"
format: html
editor: visual
---

# Task 1: Basic Vector Practice

## 1.

First, we create two vectors corresponding to the blood pressure of 20 subjects. The first, `pre_bp` is pre-treatment measurements, and the second, `post_bp` is post-treatment measurements.

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 
            128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119,
             130, 122, 106, 106, 124, 102, 117, 113)
```

## 2.

Next, we will assign `names` to the vector elements using this `paste()` function

```{r}
# Initialize vector of names
patient <- paste("Subject", 1:20, sep = "_")
patient
```

```{r}
# Update vector attributes
names(pre_bp) <- patient
names(post_bp) <- patient

# Confirm change
print(head(pre_bp))
print(head(post_bp))
```

## 3.

We can calculate the change in blood pressure by subtracting post-treatment measurements from pre-treatment measurements.

- Note: Specified as subtracting post-treatment from pre-treatment. Therefore, a positive value of `diff_bp` is a decrease in blood pressure post-treatment (negative = increase).

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

## 4.

Now, we can calculate the average decrease in blood pressure across all patients

```{r}
mean(diff_bp)
```

From this data, blood pressure decreased 17 bpm, on average, after treatment.

## 5.

We can specifically determine which patients experienced a decrease in blood pressure after treatment (positive change) using the `which()` function

```{r}
is_pos <- which(diff_bp > 0)
is_pos # values are the indeces, names are preserved
```

## 6.

We can see that all patients experienced a decrease in blood pressure after treatment except for subject 9, 13, and 17. Therefore, we can make a vector that is a subset of the differences to only include subjects whose blood pressure decreased.

```{r}
positive_diff_bp <- diff_bp[is_pos]
print(positive_diff_bp)
```

## 7.

Finally, we can recalculate the average decrease in blood pressure from this subset

```{r}
mean(positive_diff_bp)
```

From this subset of only subjects who experienced a decrease in blood pressure after treatment, the average decrease for this sample was 20.647 bpm

# Task 2: Basic Data Frame Practice

## 1.

Continuing from task 1, we'll create a data frame object with four columns corresponding to: `patient`, `pre_bp`, `post_bp`, and `diff_bp`

```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df
```

## 2.

Next, let's return only rows where `diff_bp` is negative. We'll use the`[` function for this.

Using `bp_df[ , ]`, we can subset rows (first argument) or columns. By specifying `bp_df[4] < 0` in the rows argument, we are subsetting for rows where the **fourth** column of our dataframe, `diff_bp` is less than 0.

```{r}
bp_df[bp_df[4] < 0, ]
```

## 3.

Next, let's add a new boolean column to our dataframe that returns TRUE if `post_bp` is less than 120. We can do this by using `$` to reference a column.

```{r}
bp_df$low_post <- bp_df$post_bp < 120
bp_df
```

Here we referenced the `post_bp`, column, and by giving it the argument of \< 120, this returns TRUE if the conditions are met for this column, and FALSE if not. We then can assign these boolean values to a new column in our dataframe, which I've named `low_post`

## 4.

Finally, let's print the data frame out nicely in our final document

```{r}
knitr::kable(bp_df, caption = "Blood Pressure by Subject (Treatment)",
             row.names = FALSE, 
             align = c("l", "c", "c", "c", "r") # column alignment
             )
```

# Task 3: List Practice

## 1.

Now, we have similar data from another experiment where the treatment was actually a placebo.

We'll have new data for pre and post treatment values for 10 subjects, which we'll store in a dataframe structured similarly to the dataframe from task 2

```{r}
# Initial placebo data
placebo_pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# Create difference vector
placebo_diff_bp <- placebo_pre_bp - placebo_post_bp

# Vector of placebo names, assign names to placebo pre and post
placebo_patient <- paste("Subject", 1:10, sep = "_")
names(placebo_pre_bp) <- placebo_patient
names(placebo_post_bp) <- placebo_patient

```

```{r}
# Combine as dataframe
placebo_bp_df <- data.frame(placebo_patient, placebo_pre_bp, placebo_post_bp, placebo_diff_bp)

# Add column for flagging low post bp
placebo_bp_df$low_post <- placebo_bp_df$placebo_post_bp < 120

knitr::kable(placebo_bp_df, caption = "Blood Pressure by Subject (Placebo)",
             row.names = FALSE, 
             align = c("l", "c", "c", "c", "r") # column alignment
             )
```

## 2.

Now, we can create and store a list with two elements:

- 1st element: `treatment`, our data frame from task 2
- 2nd element: `placebo`, our data frame created above

```{r}
my_list <- list(treatment = bp_df, placebo = placebo_bp_df)

# Look at list structure
str(my_list)
```

## 3.

We can access the first list element in three different ways:

First, using single brackets

```{r}
my_list[1] # by index, returned as a list
```

Second, using double brackets

```{r}
my_list[[1]] # by index, returned at itself (dataframe)
```

And third, using `$`

```{r}
my_list$treatment # by name, returned as itself
```

## 4.

Finally, let's access the `pre_bp` column from the `placebo` dataframe. We can do this via multiple bracket subsets.

```{r}
my_list[[2]][2] # list within a list, pre_bp is column 2 of our dataframe
```

# Task 4: Control Flow Practice

## 1.

Working *within* the list, let's characterize the post-treatment (or placebo) bp measurement as one of the following: \* Optimal (\<=120) \* Borderline (120 \< bp \<= 130) \* High (\>130)

We can do this by first creating a new `status` column

```{r}
my_list$treatment$status <- character(20) # Create a new empty character column for treatment

my_list$placebo$status <- character(10) # For placebo
```

## 2.

Now, we'll use a loop with if/then/else logic to fill this column

```{r}
for (i in 1:length(my_list$treatment$post_bp)) { #Reference my_list, not the original object
  if (my_list$treatment$post_bp[i] <= 120) {
    my_list$treatment$status[i] <- "Optimal"   # Assign values to the empty character column within the loop
  } else if (my_list$treatment$post_bp[i] <= 130) {
    my_list$treatment$status[i] <- "Borderline"
  } else if (my_list$treatment$post_bp[i] > 130){
    my_list$treatment$status[i] <- "High"
  } else {
    cat("Something broke at", i) # Error tracking if nothing fits these conditions 
                                #(e.g a non-number is passed into the loop)
  }
}
```

## 3.

Now, we can do the same for the placebo dataframe in our list. Here we create a new empty character column with 10 observations instead of 20.

```{r}
for (i in 1:length(my_list$placebo$placebo_post_bp)) {
  if (my_list$placebo$placebo_post_bp[i] <= 120) {
    my_list$placebo$status[i] <- "Optimal"
  } else if (my_list$placebo$placebo_post_bp[i] <= 130) {
    my_list$placebo$status[i] <- "Borderline"
  } else if (my_list$placebo$placebo_post_bp[i] > 130){
    my_list$placebo$status[i] <- "High"
  } else {
    cat("Something broke at", i) # Error tracking
  }
}
```

Finally, let's view the results

- Treatment group

```{r}
my_list$treatment # Review results
```

- Placebo group

```{r}
my_list$placebo # Review results
```
