---
title: "Homework 2"
author: "Kayla Nowak"
format: html
editor: visual
---

## Task 1: Basic Vector practice

1.  Create pre-treatment vector and post-treatment vector.

```{r}
bp_pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

bp_post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Assign (subject) names to the vector elements.

```{r}
#Initialize vector of names as "Subject_#"
sbj <- paste("Subject", 1:20, sep="_")
sbj

#Update vector attributes
names(bp_pre)  <- sbj
names(bp_post) <- sbj

#Confirm update
names(bp_pre)
names(bp_post)
```

3.  Calculate change in blood pressure.

    -   Note: Specified as subtracting post-treatment *from* pre-treatment. Therefore, a positive value of `bp_chg` is a decrease in blood pressure post-treatment (negative = increase).

```{r}
bp_chg <- bp_pre - bp_post
bp_chg 
```

4.  Average decrease across all participants.

```{r}
mean(bp_chg)   #positive == decrease
```

5.  Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
is_pos <- which(bp_chg > 0)
is_pos  #values are the indices, names are preserved (matches index)
```

6.  Subset the vector of differences to only return those that have a positive change.

```{r}
bp_chg_pos <- bp_chg[is_pos]
bp_chg_pos
```

7.  Average decrease across those with decrease.

```{r}
mean(bp_chg_pos)
```

## Task 2: Basic Data Frame practice

1.  Create data frame from the bp vectors, rename columns.

```{r}
bp_df <- data.frame(patient = sbj, 
                    pre_bp  = bp_pre, 
                    post_bp = bp_post, 
                    diff_bp = bp_chg)
```

2.  Return only rows with negative change (increase bp). **?moving forward or just for now?**

```{r}
subset(bp_df, diff_bp < 0)
```

3.  Add logical indicator column for post-treatment bp \< 120.

```{r}
bp_df$post_bp_lt120 <- bp_df$post_bp < 120
bp_df
```

4.  Print data frame (formatted with knitr).

```{r}
knitr::kable(bp_df, 
             row.names = FALSE,  #turn off row.names, redundant
             align = c("l", "r", "r", "r", "r") #force last col to r align
             )
```

## Task 3: List practice

```{r}

```

## Task 4: Control Flow Practice

```{r}

```
