---
title: "Homework 2"
author: "Kayla Nowak"
format: html
editor: visual
---

## Task 1: Basic Vector practice

1.  Create pre-treatment vector and post-treatment vector.

```{r}
bp_pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

bp_post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Assign (subject) names to the vector elements.

```{r}
#Initialize vector of names as "Subject_#"
sbj <- paste("Subject", 1:20, sep="_")
sbj

#Update vector attributes
names(bp_pre)  <- sbj
names(bp_post) <- sbj

#Confirm update
names(bp_pre)
names(bp_post)
```

3.  Calculate change in blood pressure.

    -   Note: Specified as subtracting post-treatment *from* pre-treatment. Therefore, a positive value of `bp_chg` is a decrease in blood pressure post-treatment (negative = increase).

```{r}
bp_chg <- bp_pre - bp_post
bp_chg 
```

4.  Average decrease across all participants.

```{r}
mean(bp_chg)   #positive == decrease
```

5.  Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
is_pos <- which(bp_chg > 0)
is_pos  #values are the indices, names are preserved (matches index)
```

6.  Subset the vector of differences to only return those that have a positive change.

```{r}
bp_chg_pos <- bp_chg[is_pos] #since is_pos is a vector of the indices
bp_chg_pos
```

7.  Average decrease across those with decrease.

```{r}
mean(bp_chg_pos)
```

## Task 2: Basic Data Frame practice

1.  Create data frame from the bp vectors, rename columns.

```{r}
bp_df <- data.frame(patient = sbj, 
                    pre_bp  = bp_pre, 
                    post_bp = bp_post, 
                    diff_bp = bp_chg)
```

2.  Return only rows with negative change (increase bp).

    - Note: I assumed this was asking for a printing of select rows and not a permanent subset/reduction of the data frame.

```{r}
subset(bp_df, diff_bp < 0)
```

3.  Add logical indicator column for post-treatment bp \< 120.

```{r}
bp_df$post_bp_lt120 <- bp_df$post_bp < 120
bp_df
```

4.  Print data frame (formatted with knitr).

```{r}
knitr::kable(bp_df, 
             row.names = FALSE,  #turn off row.names, redundant
             align = c("l", "r", "r", "r", "r") #force last col to r align
             )
```

## Task 3: List practice

Now working with the placebo group (previous data was the treatment group).

1.  Create a bp data frame for the placebo group (same process as before).

```{r}
#initiate placebo data (pre and post)
pl_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
pl_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

#derive change for placebo
pl_chg <- pl_pre - pl_post

#flag post_bp < 120 for placebo 
pl_flag <- pl_post < 120

#add subject identifiers (1:10, but different 1:10 than TRT group)
sbj_pl <- paste("Subject", 1:10, sep="_")
names(pl_pre)  <- sbj_pl
names(pl_post) <- sbj_pl

#combine as data frame:
pl_df <- data.frame(patient = sbj_pl, 
                    pre_bp  = pl_pre, 
                    post_bp = pl_post, 
                    diff_bp = pl_chg,
                    post_bp_lt120 =pl_flag)
pl_df

```

2.  Create list containing `treatment` data frame and `placebo` dataframe.

```{r}
bp_list <- list(treatment = bp_df,
                placebo = pl_df)
str(bp_list)
```

3.  Access the first list element using 3 different types of syntax

    -   Note: all 3 outputs should look nearly identical

```{r}
bp_list[1]          #by index, returned as list
bp_list[[1]]        #by index, returned as itself (data frame)
bp_list$treatment   #by name,  returned as itself (data frame)
```

4.  Access `pre_bp` column of placebo data in one line of code.

```{r}
bp_list[[2]][2]    # [[list]][within list], pre_bp is column 2 of placebo data frame
```

## Task 4: Control Flow Practice

Working *within* the list, add a new field to each data frame to categorize `post_bp` to one of the following levels:

-   optimal (\<= 120)
-   borderline (\>120 but \<= 130)
-   high (\> 130)

1.  Initialize new `status` column in each dataset

```{r}
#update df WITHIN each list
bp_list[[1]]$status <- character(20)  #treatment dataframe, blank for 20 records
bp_list[[2]]$status <- character(10)  #placebo dataframe, blank for 10 records
```

2.  Apply categories for the non-placebo group:

```{r}
#For loop indexes over subject (row number)
# - bp_list[[1]] isolates the treatment df list element as a df
# - [i,3] is ith row (subject) and 3rd column (post_bp)
# - [i,6] is ith row (subject) and 6th column (status)
# - logical check on 3rd column dictates the value of the 6th column

for (i in 1:20) {
  if (bp_list[[1]][i,3] <= 120){
    bp_list[[1]][i,6] = "optimal"
  }
  else if (bp_list[[1]][i,3] <= 130) {
    #Note: >120 is implied by the if/else-if logic flow
    bp_list[[1]][i,6] = "borderline"
  }
  else if (bp_list[[1]][i,3] > 130) {
    bp_list[[1]][i,6] = "high"
  } 
  else {
    cat("Something broke at ", i)
  }
}

#review results after both loops finish executing to check for indexing
```

3.  Apply categories for the placebo group:

```{r}
#Same approach, except...
# - loop over 10 subjects (instead of 20)
# - bp_list[[2]] isolates the placebo df list element as a df

for (i in 1:10) {
  if (bp_list[[2]][i,3] <= 120){
    bp_list[[2]][i,6] = "optimal"
  }
  else if (bp_list[[2]][i,3] <= 130) {
    #else-if logic implies also (not <= 120)
    bp_list[[2]][i,6] = "borderline"
  }
  else if (bp_list[[2]][i,3] > 130) {
    bp_list[[2]][i,6] = "high"
  } 
  else {
    cat("Something broke at ", i)
  }
}

```

Review results:

- Non-Placebo group

```{r}
bp_list[[1]]
```

- Placebo group

```{r}
bp_list[[2]]
```

