---
title: "ST558_HW3_Kayla"
author: "Kayla Nowak"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  **If your working directory is `myfolder/homework/`, what *relative* path would you specify to get the file located at `myfolder/MyData.csv`?**

    Since the file we want is only one folder above the current working directory, we can simply specify `../MyData.csv` as the relative file path. The `../` path prefix points to one folder above the current working directory.

2.  **What are the major benefits of using R projects?**

    R projects support the use of working directories and version control software. Working directories establish relative paths, which eliminate the need for each user to hard-code file paths specific to their machine. Relative paths are very helpful for code sharing, and version control is also helpful for code sharing. Through an R project for version control we can clone a Git repository and manage that repository through R Studio's terminal.

3.  **What is `Git` and what is `GitHub`?**

    Git a type of version control software. GitHub is an online platform for hosting and sharing Git-based projects (the actual version control actions still happen within Git).

4.  **What are the two main differences between a `tibble` and a `data.frame`?**

    1)  Fancy print. A data frame prints in its entirety vs. a tibble prints the first few rows of data and a size summary (#rows x #cols).
    2)  Limited type coercion. A single column of a tibble can be accessed by `[` *without* coercing to a vector.

# Task 2: Reading Delimited Data

For both data sets, directly read the data into R directly from the URL using functions from the tidyverse. Add column names and print out the tibble.

## 1.

The Glass data file is comma delimited and has 11 columns:

-   column 1 is a numeric ID value (1-214)
-   columns 2:10 are numeric measurements
-   column 11 is a numeric code for type of glass

Source: <https://www4.stat.ncsu.edu/online/datasets/glass.data>

```{r}
library(readr)

#read from URL, as tibble
glass <- read_delim("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", 
                                  "K", "Ca", "Ba", "Fe", "TypeOfGlass"))

#print for review
glass
```

The default data types look good.

## 2.

The Yeast data file has a fixed-width format with 10 columns:

-   column 1 is an alphanumeric ID value
-   columns 2:9 are numeric measurements
-   column 10 is a character site ID

Source: <https://www4.stat.ncsu.edu/online/datasets/yeast.data>

```{r}
#read from URL, as tibble
yeast <- read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                  fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                  c("seq_name", "mcg", "gvh", "alm", "mit",
                                "erl", "pox", "vac", "nuc", "class")))

#print for review
yeast
```

The default data types look good.

# Task 3: Combining Excel and Delimited Data

Working with the Portuguese “Vinho Verde” wine data.

## 3.

Import white wine data from excel.

Source: [https://www4.stat.ncsu.edu/online/ datasets/white-wine.xlsx](https://www4.stat.ncsu.edu/online/datasets/white-wine.xlsx)

```{r}
library("readxl")

#read data (Sheet 1), saved in repo/HW3 folder (same as .qmd)
white_wine <- read_xlsx(path="white-wine.xlsx",
                        sheet="white-wine")

#print
white_wine
```

## 4.

Convert to standard variable names (found on Sheet 2).

```{r}

#read data (Sheet 2), saved in repo/HW3 folder (same as .qmd)
white_wine_vars <- read_xlsx(path="white-wine.xlsx",
                             sheet="variables")

#view structure/contents
str(white_wine_vars)

#update col names on white_wine
colnames(white_wine) <- white_wine_vars$Variables

#confirm update
white_wine

```

## 5.

Add a column identifying all records as white wine.

```{r}
white_wine$wine_type <- "white"

#confirm update
xtabs(~wine_type, white_wine)

```

# Task 4: Writing a Function
