---
title: "ST558 Homework 3"
author: "Ryan Mersereau"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

    We can use `../MyData.csv`, typically inside a `read.csv()` function. The `..` prefix means we go up one level from our current working directory, putting us inside `myfolder`

2.  What are the major benefits of using R projects?

    When an R project is created, it gets associated with a directory (folder or repo) and uses this as the projects working directory. This is very useful, because if we have this repo shared (via Github), other users can pick up the same R project and all the paths should work. This allows for organization and seamless collaboration.

3.  What is Git and what is GitHub?

    Git is the actual software that performs version contolling. whereas Github is the online hosting service for projects that utilize Git software. Version controlling allows us to save every change that occurs in our project folder, and revert back to previous saves.

4.  What are the two main differences between a tibble and a dataframe?

    Tibbles are a modern re-imagining of the data frame. When printing a tibble, we only see the first few rows/columns, whereas printing a dataframe prints everything to the console. Tibbles also force you to confront problems earlier when subsetting, as they dont do partial matching. `df$colnam` on a dataframe will match a column named `colname`, whereas with a tibble it warns you explicitly the column doesn't exist

# Task 2: Reading Delimited Data

## 1. Glass Data

The first data set we'll work with is called `glass.data`, from the UCI machine learning repository. <https://www4.stat.ncsu.edu/~online/datasets/glass.data.>

We'll read this data into R directly from the URL using `tidyverse` functions. The delimiter for this data are commas, so we will specify that with the `read_csv` function.

```{r}
#| message: False
#| warning: False

library(tidyverse)
```

```{r}
# Read in the file using the read_csv() function from the tidyverse `readr` package

glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                   "Ca", "Ba", "Fe", "Type"))
# Print tibble
glass
```

## 2. Yeast Data

<https://www4.stat.ncsu.edu/~online/datasets/yeast.data.>

The second is called `yeast.data`. We'll also read this into R from a url using `tidyverse` functions. This data is white space delimited, so we can use `read_table()` from the `readr` package

```{r}
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
    col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac",
             "nuc", "class"))

# Print tibble
yeast
```

# Task 3: Combining Excel and Delimited Data

For this step, we'll use two datasets related to red and white wine variants. This data has physicochemical variables and a quality score.

## 3. White wine data

<https://www4.stat.ncsu.edu/~online/datasets/white-wine.xlsx.>

First we'll access an excel version of the white wine dataset placed in the HW3 folder (where this .qmd lives), and read it in using the `readxl` package.

```{r}
#| message: False
#| warning: False

library(readxl)
```

```{r}
# Reads first sheet by default
white_wine <- read_excel("white-wine.xlsx")

# Print out tibble
white_wine
```

## 4. Variable name standardization

We notice from printing white_wine to the console that the variable names are surrounded by backticks. This is because they are non-standard (include a space). Let's fix this by reading in the variable names on the second sheet of the excel file

```{r}
# Read variable names from sheet 2 (named variables)
white_wine_variables <- read_excel("white-wine.xlsx", sheet = "variables")

# Re-assign column names
colnames(white_wine) <- white_wine_variables$Variables

# Print tibble
white_wine
```

## 5. Color indicator column

Now, let's add a column to this data set indicating these wines are white (they all are) using the `$` operator

```{r}
# Add column, all of them are white wines so this is the only option
white_wine$type <- "white"

# Check tibble once again
white_wine
```

## 6. Red wine data

Now, there is a semi-colon delimited file of red wine data we can read in with the url below. 

<https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv.>

We'll do this with the `read_delim()` function from `readr`, specifying the delimiter type

```{r}
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim=";")

red_wine
```

We'll do the same modifications we did for our white wine data: replacing variable names and appending a `type` column, this time denoting red for all observations

```{r}
# Rename columns
colnames(red_wine) <- white_wine_variables$Variables

# Add type column
red_wine$type <- 'red'

# Print tibble
red_wine
```

## 7. Combining data 

Finally, let's combine the white and red wine data sets into one! We can use `dplyr::bind_rows()`
to create a single tibble containing all the wine data

```{r}
# Create total wine data sets
total_wine <- dplyr::bind_rows(white_wine, red_wine)

# Print
total_wine
```

# Task 4: Writing a Function

## 8. IQR function

We start this task by writing a basic function to get the IQR from a vector of numbers using the `quantile` function

```{r}
getIQR <- function(vector, ...) { # ... Allows other unnamed arguments to be passed
  
  q <- quantile(vector, probs = c(0.25, 0.75), names = F, ...) 
  # Saving this as q allows us to only have to use `quantile` once
  return(q[2] - q[1])
}
```

## 9. Applying IQR function

Now, lets take the total_wine dataset and apply the `getIQR` function to `fixed_acidity` and `citric_acid`

```{r}
getIQR(total_wine$fixed_acidity)
```
The IQR for fixed acidity is 1.3

```{r}
getIQR(total_wine$citric_acid)
```
The IQR for citric acid is 0.14

## 10. Coefficient of variation function

Let's write another function that returns the coefficient of variation of a vector, defined as the sample standard deviation divided by the absolute value of the sample mean.

```{r}
getCoefVar <- function(vector, ...){
  return((sd(vector, ...)) / abs(mean(vector, ...)))
  }
```



## 11. Applying CoefVar function

Same with step 9, we'll apply our newly created function to the two total wine columns

```{r}
getCoefVar(total_wine$fixed_acidity)
```
The coefficient of variation for fixed acidity is 0.1796

```{r}
getCoefVar(total_wine$citric_acid)
```

The coefficient of variation for citric acid is 0.456

## 12. Wrapper function

Now, let's write a function `getSummaries()` that takes in a vector and returns both the 
IQR and the coefficient of variation by calling both of these previous functions

```{r}
getSummaries <- function(vector, ...){
  
  return(list(IQR = getIQR(vector, ...), Coef_Var = getCoefVar(vector, ...)))
}
```


## 13. Applying wrapper function

Let's use our new wrapper function on the two columns in total_wine again
```{r}
getSummaries(total_wine$fixed_acidity)
```

```{r}
getSummaries(total_wine$citric_acid)
```

## 14. Infix function

Finally, we'll write an **infix** function that can be used in place of the `cbind` function. We'll check that this works with some sample dataframes and a vector.

```{r}
# Sample dataframes and vector
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

# View objects
df1
df2
vec1
```

```{r}
# Create function
`%cb%` <- function(x, y){
  cbind(x, y)
}
```

First, lets test with both dataframes
```{r}
df1 %cb% df2
```

Finally, test with both dataframes and vector
```{r}
df1 %cb% df2 %cb% vec1
```

Creating this infix function lets us cleanly bind these arguments together in one line without the added function notation

