---
title: "ST558_HW3_Kayla"
author: "Kayla Nowak"
format: html
editor: visual
---

# Task 1: Conceptual Questions

1.  **If your working directory is `myfolder/homework/`, what *relative* path would you specify to get the file located at `myfolder/MyData.csv`?**

    Since the file we want is only one folder above the current working directory, we can simply specify `../MyData.csv` as the relative file path. The `../` path prefix points to one folder above the current working directory.

2.  **What are the major benefits of using R projects?**

    R projects support the use of working directories and version control software. Working directories establish relative paths, which eliminate the need for each user to hard-code file paths specific to their machine. Relative paths are very helpful for code sharing, and version control is also helpful for code sharing. Through an R project for version control we can clone a Git repository and manage that repository through R Studio's terminal.

3.  **What is `Git` and what is `GitHub`?**

    Git a type of version control software. GitHub is an online platform for hosting and sharing Git-based projects (the actual version control actions still happen within Git).

4.  **What are the two main differences between a `tibble` and a `data.frame`?**

    1)  Fancy print. A data frame prints in its entirety vs. a tibble prints the first few rows of data and a size summary (#rows x #cols).
    2)  Limited type coercion. A single column of a tibble can be accessed by `[` *without* coercing to a vector.

# Task 2: Reading Delimited Data

For both data sets, directly read the data into R directly from the URL using functions from the tidyverse. Add column names and print out the tibble.

## 1.

The Glass data file is comma delimited and has 11 columns:

-   column 1 is a numeric ID value (1-214)
-   columns 2:10 are numeric measurements
-   column 11 is a numeric code for type of glass

Source: <https://www4.stat.ncsu.edu/online/datasets/glass.data>

```{r}
library(readr)

#read from URL, as tibble
glass <- read_delim("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", 
                                  "K", "Ca", "Ba", "Fe", "TypeOfGlass"))

#print for review
glass
```

The default data types look good.

## 2.

The Yeast data file has a fixed-width format with 10 columns:

-   column 1 is an alphanumeric ID value
-   columns 2:9 are numeric measurements
-   column 10 is a character site ID

Source: <https://www4.stat.ncsu.edu/online/datasets/yeast.data>

```{r}
#read from URL, as tibble
yeast <- read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                  fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                  c("seq_name", "mcg", "gvh", "alm", "mit",
                                "erl", "pox", "vac", "nuc", "class")))

#print for review
yeast
```

The default data types look good.

# Task 3: Combining Excel and Delimited Data

Working with the Portuguese “Vinho Verde” wine data.

## 3.

Import white wine data from excel.

Source: [https://www4.stat.ncsu.edu/online/ datasets/white-wine.xlsx](https://www4.stat.ncsu.edu/online/datasets/white-wine.xlsx)

```{r}
library("readxl")

#read data (Sheet 1), saved in repo/HW3 folder (same as .qmd)
white_wine <- read_xlsx(path="white-wine.xlsx",
                        sheet="white-wine")

#print
white_wine
```

## 4.

Convert to standard variable names (found on Sheet 2).

```{r}

#read data (Sheet 2), saved in repo/HW3 folder (same as .qmd)
white_wine_vars <- read_xlsx(path="white-wine.xlsx",
                             sheet="variables")

#view structure/contents
str(white_wine_vars)

#update col names on white_wine
colnames(white_wine) <- white_wine_vars$Variables

#confirm update
white_wine

```

## 5.

Add a column identifying all records as white wine.

```{r}
white_wine$type <- "white"

#confirm update
xtabs(~type, white_wine)

```

## 6.

Import red wine data from excel (semicolon delimited csv). Again, clean variable names and add wine type identifier.

Source: [https://www4.stat.ncsu. edu/online/datasets/red-wine.csv](https://www4.stat.ncsu.edu/online/datasets/red-wine.csv)

```{r}
library("readr")

#read data, saved in repo/HW3 folder (same as .qmd)
red_wine <- read_delim(file="red-wine.csv",
                       delim=";")

#print - confirm all doubles
red_wine

#confirm columns/order are the same in both sources
colnames(white_wine)
colnames(red_wine)

#update column names
colnames(red_wine) <- white_wine_vars$Variables 

#add type column
red_wine$type <- "red"

#confirm update
xtabs(~type, red_wine)
```

## 7.

Combine white and red wine data in a single data set.

```{r}

#row bind since shared structure
all_wine <- dplyr::bind_rows(white_wine, red_wine)

#review
all_wine
xtabs(~type, all_wine)

```

# Task 4: Writing a Function

## 8.

Write a function that returns the interquartile range (P75 - P25) for a numeric vector input.

```{r}

getIQR <- function(x) {
  
  if(!is.vector(x) | !is.numeric(x)){   
    #check for valid input before proceeding
    stop("Invalid input, requires numeric vector.")
  }
  else {
    
    #q stores P25 and P75 as vector elements, names disabled
    q <- quantile(x, probs = c(0.25, 0.75), names=FALSE)
    
    #iqr captures the calculated difference:  P75 - P25
    iqr <- q[2] - q[1]
    return(iqr)
    
  }
}

## Function tests (uncomment to run):

#getIQR(c(1:10))                     #should return 4.5
#test2 <- c(1:10); getIQR(test2)     #should return 4.5

#getIQR(c("A","B"))                  #should stop (character vector)
#test3 <- c("A","B"); getIQR(test3)  #should stop (character vector) 
#getIQR(iris)                        #should stop (not a vector)

```

## 9.

Apply `getIQR()` to the `fixed_acidity` and `citric_acid` columns of the combined wine dataset from question #7.

```{r}
getIQR(all_wine$fixed_acidity)
getIQR(all_wine$citric_acid)
```

## 10.

Write a function that returns the sample coefficient of variation (sample SD / absolute(sample mean)).

```{r}

getCoefVar <- function(x){
    
  if(!is.vector(x) | !is.numeric(x)){   
    stop("Invalid input, requires numeric vector.")
  }
  else {
    coefVar <- sd(x) / abs( mean(x) )
    return(coefVar)
  }
}

## Function tests (uncomment to run):
#getCoefVar(c(1:10))                     #should return 0.55
#test4 <- c(1:10); getCoefVar(test4)     #should return 0.55

```

## 11.

Apply `getCoefVar()` to the `fixed_acidity` and `citric_acid` columns of the combined wine dataset from question #7.

```{r}
getCoefVar(all_wine$fixed_acidity)
getCoefVar(all_wine$citric_acid)
```

## 12.

Write a wrapper function that returns both IQR and coefficient of variation.

```{r}

getSummaries <- function(x){
    
  if(!is.vector(x) | !is.numeric(x)){   
    stop("Invalid input, requires numeric vector.")
  }
  else {
    
    #calculate summaries
    IQR <- getIQR(x)
    CoefVar <- getCoefVar(x)
    
    #combine, name, and return
    summaries <- c(IQR=IQR, CoefVar=CoefVar)
    return(summaries)
  }
}

## Function tests (uncomment to run):
#getSummaries(c(1:10))                     #should return IQR=4.5, CoefVar=0.55
#test5 <- c(1:10); getSummaries(test5)     #should return IQR=4.5, CoefVar=0.55

```

## 13.

Apply `getCoefVar()` to the `fixed_acidity` and `citric_acid` columns of the combined wine dataset from question #7.

```         
Note: Assuming the instructions meant to reference question 9 activities (re: question 7 data) since question 2 was about importing the yeast data.
```

```{r}
getSummaries(all_wine$fixed_acidity)
getSummaries(all_wine$citric_acid)
```
