---
title: "ST558_HW3_Yefrid"
author: "Yefrid Cordoba"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Task 1: Conceptual questions

1.  If my working directory is `myfolder/homework/`, and I need to reference a file located at `myfolder/MyData.csv`, I need to go one level above `(using ..)`my working directory followed by the file name in this order: `../MyData.csv`

2.  The primary benefit of using `R` projects is the ease to keeping different workflows separate, ensuring that each project maintains its own working directory, workspace, history, folder structure, and source documents.

3.  Git is an open source version control software which uses snapshots of the work to keep track of changes rather than creating multiple copies of the files.\
    Github is the cloud base platform that host Git repositories, from which you can collaborate and share your repository.

4.  A tibble has multiple classes and one of that is a `data.frame`, which is open when methods apply for different classes, being first `tbl_df`, then `tbl`, and finally `data.frame`.\
    They are very similar in functionality, however, when the method `print()` is called, we get more information of the tibble like number of rows, and columns, and are actually are printed just the first few.

## Task 2: Reading delimited data

### 1.

First we load the `tidyverse` library

```{r}
#| message: False
#| warning: False

library(tidyverse)
```

And now we read the file, adding column names, from the link, the data is separated by comma, therefore, we use `read_csv()` from the `tidyverse`

```{r}
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                     "Ca", "Ba", "Fe", "Type"))
glass_data
```

### 2.

Based on the link, the delimiter for this data set is a double space, therefore we need to use the `read_delim()` to read this dataset.
```{r}
#first we create the vector of names
n_yeast <- c("seq_name", "mch", "gvh", "alm", "mit", "erl", "pox", "vac",
             "nuc", "class")

#and now we create the tibble
yeast_data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                       delim = "  ",
                       col_names = n_yeast)
yeast_data
```

## Task 3: Combining Excel and delimited data

First we import the package `readxl`
```{r}
#| message: False
#| warning: False

library(readxl)
```

### 3.

And now we read the data from the first sheet of the excel file

```{r}
#by default reads from the first sheet of the document
white_data <- read_excel("white-wine.xlsx")

white_data
```

### 4.

Now we read just the first column of names from the sheet two to get the tibble with variable names.

```{r}
var_wine <- read_excel("white-wine.xlsx",
                       sheet = excel_sheets("white-wine.xlsx")[2])
#And we re-assign the colnames to the tibble with the data

colnames(white_data) <- var_wine$Variables #it is necessary to get the varaible column as a vector of characters

white_data

```

### 5. 

And we add a column `type` for all the rows of this tibble using the `$` operator

```{r}
white_data$type <- "white"

white_data
```

### 6.

Now we need to read the red wine data set
```{r}
red_data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";",
                       col_names = var_wine$Variables,
                       skip = 1
                      )
red_data
```

And add the column of `red` for all the observations

```{r}
red_data$type <- "red"

red_data
```

### 7. 

Finally we combine the two tibbles

```{r}
wine_data <- dplyr::bind_rows(white_data, red_data)
wine_data
```

## Task 4: Writing a function

### 8.

We are going to write a function to get the interquartile range of a vector, we use `names = F` to remove the percentages names. And add `...` to allow unamed arguments for the quantile function.
```{r}
getIQR <- function(v, ...){
  
  quantile(v, names = F, ...)[4] - quantile(v, names = F, ...)[2]
  
  
}

```

### 9.

Now, we are going to apply the `getIQR()` function to the `fixed_acidity`, 

```{r}
getIQR(wine_data$fixed_acidity)
```
and the `citric_acid columns`.
```{r}
getIQR(wine_data$citric_acid)
```

### 10.

We are going to write a function to get the coefficient of variation

```{r}
getCoefVar <- function(v, ...){
  
  sd(v, ...) / abs(mean(v, ...))
  
}
```

### 11.

And we apply to the `fixed_acidity` column,

```{r}
getCoefVar(wine_data$fixed_acidity)
```
and to the `citric_acid` column.

```{r}
getCoefVar(wine_data$citric_acid)
```

### 12.

We need to create a wrapper function that return the IQR and the coefficient of variation for a given vector. The output is a list with both of the values calculated.

```{r}
getSummaries <- function(v, ...){
  
  list(IQR = getIQR(v, ...), Coef_Var = getCoefVar(v, ...))
  
}
```

### 13.

We apply the new functio to the `fixed_acidity`
```{r}
getSummaries(wine_data$fixed_acidity)
```

And to the `citric_acid` columns.

```{r}
getSummaries(wine_data$citric_acid)
```

### 14.

We need to write an infix function that replace the `cbind` function.

```{r}
`%cb%` <- function(v1, v2){
  
  cbind(v1, v2)
  
}
```

We generate some `df` to apply the infix function

```{r}
set.seed(12)
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
```

And finally we use the infix function with 2 data frames

```{r}
df1 %cb% df2
```

And with 2 data frames and 1 vector

```{r}
df1 %cb% df2 %cb% vec1
```

