---
title: "ST558_HW3_Yefrid"
author: "Yefrid Cordoba"
format: html
editor: visual
---

## Task 1: Conceptual questions

1.  If my working directory is `myfolder/homework/`, and I need to reference a file located at `myfolder/MyData.csv`, I need to go one level above `(using ..)`my working directory followed by the file name in this order: `../MyData.csv`

2.  The primary benefit of using `R` projects is the ease to keeping different workflows separate, ensuring that each project maintains its own working directory, workspace, history, folder structure, and source documents.

3.  Git is an open source version control software which uses snapshots of the work to keep track of changes rather than creating multiple copies of the files.\
    Github is the cloud base platform that host Git repositories, from which you can collaborate and share your repository.

4.  A tibble has multiple classes and one of that is a `data.frame`, which is open when methods apply for different classes, being first `tbl_df`, then `tbl`, and finally `data.frame`.\
    They are very similar in functionality, however, when the method `print()` is called, we get more information of the tibble like number of rows, and columns, and are actually are printed just the first few.

## Task 2: Reading delimited data

First we load the `tidyverse` library

```{r}
#| show: False

library(tidyverse)
```
And now we read the file, adding column names, from the link, the data is separated by comma, therefore, we use `read_csv()` from the `tidyverse`

```{r}
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K",
                                     "Ca", "Ba", "Fe", "Type"))
glass_data
```

