---
title: "ST558_HW5_Yefrid"
author: "Yefrid Cordoba"
format: html
---

## Task 1: Querying an API and a Tidy-Style Function

### 1.

We need to use the `GET()` function from the topic trade

```{r}
#| message: False
#| warning: False

library(httr)
library(jsonlite)
library(tidyverse)
library(readr)
```

Now we obtain the info from the API.

```{r}

info_news <- GET("https://newsapi.org/v2/everything?q=trade&from=2026-08-21&language=en&pageSize=50&apiKey=b83cc47afb8e4f47a5ff9dafee63d05c")
```

We explore the info to determine the actual data that we need to extract.

```{r}
str(info_news)
```

### 2.

We see that the data is under the name `content`.   
We use the `jsonlite` package to get a tibble.

```{r}
parsed <- fromJSON(rawToChar(info_news$content))
team_info <- as_tibble(parsed$articles)
team_info
```

### 3.

Now we will create a funstion that will create the URL to get data from the `News` API.

```{r}
get_news <- function(title, period, key){
  url <- paste(paste0("https://newsapi.org/v2/everything?q=", title),
               paste0("from=", as.character(period)),
               paste0("apiKey=", key),
               sep = "&"
               )
  
  
  info_news <- GET(url = url)
  
  parsed <- fromJSON(rawToChar(info_news$content))
  
  team_info <- as_tibble(parsed$articles)
  
  team_info
  
}
```

Now we can test the function with two different topics and dates

```{r}
get_news("macklemore", "2026-09-15", "b83cc47afb8e4f47a5ff9dafee63d05c")
```

```{r}
get_news("gas", "2026-09-10", "b83cc47afb8e4f47a5ff9dafee63d05c")
```

## Task 2: Summarizing Roller Coaster Data

First we are going to read the data from the roller coaster, and convert the required variables to factor.

```{r}
rc_data <- read.csv("635USrollercoasters.csv") |>
  mutate(State = as.factor(State),
         Census.Region = as.factor(Census.Region),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         Inversions. = as.factor(Inversions.),
         Make = as.factor(Make),
         Latitude = NULL,
         Longitude = NULL,
         Boundary = NULL)

```


```{r}
rc_data
```

