---
title: "ST558_HW5_Will"
author: "Will Carter"
format: html
editor: visual
---




# Task 1: Querying an API and a Tidy-Style Function

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
```

## 1. GET() call for an interesting topic in news
```{r}
results = GET("https://newsapi.org/v2/everything?q=Iran&from=2026-08-21&language=en&pageSize=50&apiKey=8bc29bd8879e49e9940d1a64d5532623")
```

## 2. parse results for actual article
```{r}
parsed_results <- fromJSON(content(results, as = "text", encoding = "UTF-8"))

articles <- parsed_results$articles
```

```{r}
articles
```

## 3. function to easily query API
```{r}
get_news_article <- function(query, from_date, api_key) {
  url <- paste0("https://newsapi.org/v2/everything?q=", URLencode(query),
  "&from=", from_date,
  "&language=en&pageSize=50&apiKey=", api_key)
  
  result <- GET(url)
  
  parsed_result <- fromJSON(content(result, as = "text", encoding = "UTF-8"))
  
  parsed_result$articles
}
```

```{r}
api_key = "8bc29bd8879e49e9940d1a64d5532623"
```

## clean energy article
```{r}
articles_energy <- get_news_article("clean energy", "2026-08-21", api_key)
```

```{r}
articles_energy
```

## biotech article
```{r}
articles_biotech <- get_news_article("biotech", "2026-08-21", api_key)
```

```{r}
articles_biotech
```


# Task 2: Summarizing Roller Coaster Data

## 4. Look at the data
```{r}
roller_coasters <- read_csv("../HW5/635USrollercoasters.csv")
```

```{r}
glimpse(roller_coasters)
summary(roller_coasters)
```

```{r}
roller_coasters <- roller_coasters |>
    mutate( 
      AgeGroup = factor(AgeGroup, levels = c("1:Older", "2:Recent", "3:Newest"),
                        labels = c("Older", "Recent", "Newest"),
                        ordered = TRUE)
    )
```


## 5. Document any missing values in data
```{r}
colSums(is.na(roller_coasters))
```


## 6. Create one-way, two-way, and three-way contingency table
```{r}
one_way_table <- table(roller_coasters$Type)
```

```{r}
one_way_table
```

```{r}
two_way_table <- table(roller_coasters$Type, roller_coasters$Scale)
```

```{r}
two_way_table
```

```{r}
three_way_table <- table(roller_coasters$Type, roller_coasters$Scale, roller_coasters$AgeGroup)
```

```{r}
three_way_table
```


## 7. Create conditional two-way table

### using filter()
```{r}
new_coasters <- roller_coasters |> filter(AgeGroup == "Newest")
table(new_coasters$Type, new_coasters$Scale)
```

### build then subset
```{r}
three_way_table[, , "Newest"]
```

## 8. Create two-way contingency table using dplyr
```{r}
roller_coasters |> 
  group_by(Type, Scale) |> 
  summarize(n = n(), .groups = "drop") |>
  pivot_wider(names_from = Scale, values_from = n, values_fill = 0)
```

## 9. Create stacked bar graph and side by side bar graph
```{r}
ggplot(roller_coasters, aes(x = Type, fill = Scale)) + geom_bar(position = "stack") + labs(x = "Coaster Type", y = "Count", title = "Scale Distribution by Coaster Type Stacked", fill = "Scale")
```

```{r}
ggplot(roller_coasters, aes(x = Type, fill = Scale)) + geom_bar(position = "dodge") + labs(x = "Coaster Type", y = "Count", title = "Scale Distribution by Coaster Type Side-By-Side", fill = "Scale")
```



