---
title: "ST558 Homework 5 - Group 1"
author: 
  - name: Kayla Nowak
  - name: Ryan Merserau
  - name: Will Carter
  - name: Yefrid Cordoba
format: html
---

## Task 1: Querying an API in a Tidy-Style Fashion

Connect to the news API here: <https://newsapi.org/>

```{r}
#| message: False
#| warning: False

library(tidyverse)
library(httr)
library(jsonlite)

my_key <- "71ce2ed6f6844a9b8c771a2b3993de20"
```

### 1.

Use `GET()` from the `httr` package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object).

This query searches for articles from last Monday regarding Panthers quarterback Bryce Young.

```{r}
news_results <- GET(
  url = "https://newsapi.org/v2/everything",
  query = list(
    q = "bryce-young",
    from = "2026-09-14",
    language = "en",
    pageSize = 50,
    apiKey = my_key
  )
)
```

### 2.

Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
#Check top-level contents
str(news_results, max.level=1)

#Parse to contents
news_parsed <- fromJSON(rawToChar(news_results$content))
str(news_results, max.level=1)

#Select articles
articles <- as_tibble(news_parsed$articles)
articles
```

### 3.

Now write a quick function that allows the user to easily query this API.

```{r}
get_news_article <- function(subject, startdt, key) {
  
  URL <- paste0("https://newsapi.org/v2/everything?q=",
                URLencode(subject),
                "&from=", as.character(startdt),
                "&to=", as.character(Sys.Date()),
                "&language=en&pageSize=50&apiKey=", key)
  
  pulled <- GET(URL)
  #Note: missing function inputs result in R error.
    
  if(pulled$status != 200) { return("Something broke, check requirements.") }
  else {
    parsed <- fromJSON(rawToChar(pulled$content))
    tdata  <- as_tibble(parsed$articles)
    return(tdata)
  }
}
```

Test function with two different topics and dates:

```{r}
get_news_article("macklemore", "2026-09-15", my_key)
get_news_article("clean energy", "2026-09-21", my_key) #space handled by URLencode()
```

## Task 2: Summarizing Roller Coaster Data

Working with the roller coaster data. Source: <https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv>

First we are going to read the data and convert the required variables to factor. No need to use the `Boundary` variable or the `Latitude` and `Longitude` variables in the EDA.

-   Dropped ignored variables.
-   Renamed non-standard variables for ease of use.
-   Converted all categorical variables to factors.

```{r}
rollercoasters <- read_csv("635USrollercoasters.csv") |>
  select(-c(Boundary, Latitude, Longitude)) |>
  rename(CensusRegion = `Census Region`,
         YearOpened = `Year Opened`,
         Inversion_YN = `Inversions?`,
         Inversion_Num = `# of Inversions`) |> 
  mutate(State = as.factor(State),
         CensusRegion = as.factor(CensusRegion),
         AgeGroup = factor(AgeGroup, 
                           levels=c("1:Older", "2:Recent", "3:Newest"),
                           labels=c("Older", "Recent", "Newest")),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         Inversion_YN = as.factor(Inversion_YN))
```

### 4.

Look at how the data is stored and see if everything makes sense.

```{r}
#General Structure
str(rollercoasters)

#Review Factor Levels
vars_cat <- c("State", "CensusRegion", "AgeGroup", "Scale", "Type", "Design", "Inversion_YN")

for(i in 1:length(vars_cat)){
  rollercoasters |> 
    group_by(across(all_of(vars_cat[i]))) |> 
    summarize(count=n()) |> print()
}

```

-   `Coaster`, `Park`, `City`, and `Make` are fine to store as character
-   `YearOpened`, `TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, and `Inversion_Num` are appropriately stored as numeric
-   All other variables make sense to store as factors. Only `AgeGroup` benefited from a label change.

### 5.

Document any missing values in the data.

```{r}
colSums(is.na(rollercoasters))
```

`TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, and `Make` have some records with missing values. 

### 6.

Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this. Interpret a number from each table.

**One-Way Table:**

```{r}
table(rollercoasters$Type)
```

There are 530 roller coasters made of steel.

**Two-Way Table:**

```{r}
table(rollercoasters$Type, rollercoasters$Scale)
```

Out of the 530 steel roller coasters, 216 are categorized as extreme.

**Three-Way Table:**

```{r}
table(rollercoasters$Type, rollercoasters$Scale, rollercoasters$AgeGroup)
```

From the older roller coasters, 9 are made of wood and categorized as extreme

### 7.

Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods.

Looking at Scale x Design for Steel type coasters.

**By subsetting first, then making a 2-way table:**

```{r}
#single chain approach
rollercoasters |> filter(Type=="Steel") |> select(Scale, Design) |> table()
```

**By subsetting a 3-way table:**

```{r}
table(rollercoasters[, c("Scale", "Design", "Type")])[,,"Steel"]
```

### 8.

Create a two-way contingency table using `group_by()` and `summarize()` from dplyr. Then use `pivot_wider()` to make the result look more like the output from `table()`.

```{r}
rollercoasters |> 
  group_by(Type, Scale) |> 
  summarize(n = n(), .groups = "drop") |>
  pivot_wider(names_from = Scale, values_from = n, values_fill = 0)
```

### 9.

Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

**Stacked:**

```{r}
library(ggplot2)

g1 <- ggplot(data = rollercoasters |>
               drop_na(Type, Scale), aes(x = Scale, fill = Type))
g1 + geom_bar() + labs(x = "Scale", y = "Count", title = "Rollercoaster Types by Scale") 
```

**Side-by-Side:**

```{r}
ggplot(data = rollercoasters |> drop_na(Type, Scale), aes(x = Scale, fill = Type)) +
  geom_bar(position = "dodge") + 
  labs(x = "Scale", title = "Roller Coaster Types by Scale", y = "Count") +
  scale_fill_discrete("Type")
```

This is better (although the scales should be ordered). From most to least intense it goes: Extreme, Thrill, Family, Kiddie.
