---
title: "ST558 Homework 5"
author: Ryan Mersereau
format: html
editor: visual
---

# Task 1: Querying an API and a Tidy-Style Function

For this task, we'll need to connect to the news API here: <https://newsapi.org>

1. Use `GET()` from the `httr` package to return information about a topic that has been in the news 
lately

Here I chose to search for articles from last Monday regarding Panthers quarterback Bryce Young
```{r}
#| message: FALSE
#| warning: FALSE
library(httr)
library(jsonlite)
library(tidyverse)
```

```{r}
# Store personal news api key
my_key <- "48e8f55cb8e74372b534d5595a576d81"

# Building the url
id_info <- GET(
  url = "https://newsapi.org/v2/everything",
  query = list(
    q = "bryce-young",
    from = "2026-09-14",
    language = "en",
    pageSize = 50,
    apiKey = my_key
  )
)

str(id_info, max.level = 1)
```

2. Parse what is returned and find your way to the data frame that has the actual article information in it (content). The first column should be a list column!

```{r}
parsed <- fromJSON(rawToChar(id_info$content))

str(parsed, max.level = 1)
```
This returns a list with 3 elements. To specifically look at the articles published,
we can store this as a tibble below.

```{r}
parsed_articles <- as_tibble(parsed$articles)
parsed_articles
```

3. Now write a quick function that allows the user to easily query this API. The inputs
to the function should be the title/subject to search for, a time period to search from, and an
API key.

```{r}
query_function <- function(subject, date, api_key) {
  response <- GET(
    url = "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = date,
      language = "en",
      pageSize = 50,
      apiKey = api_key
    )
  )
  
  parsed <- fromJSON(rawToChar(response$content))
  as_tibble(parsed$articles)
}
```

Lets call this a couple times to show it works

```{r}
bojangles <- query_function("Bojangles", "2026-09-10", my_key)
bojangles
```

```{r}
israel <- query_function("Israel", "2026-08-31", my_key)
israel
```

# Task 2: Summarizing Roller Coaster Data

For this part, we'll use roller coaster data found here: <https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv>

Let's read in this data and convert variables to factors as needed

```{r}
#| message: FALSE
#| warning: FALSE
library(readr)
library(psych)
```


```{r}
rollercoasters <- read_csv("../HW5/635USrollercoasters.csv")
```


```{r}
rollercoasters |>
  mutate(AgeGroup = factor(AgeGroup, levels = c("1:Older", "2:Recent", "3:Newest"),
         labels = c("Older", "Recent", "Newest")),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design))
```

## Basic EDA

4. Look at how the data is stored and see if everything makes sense.

```{r}
str(rollercoasters)
```

```{r}
psych::describe(rollercoasters)
```


5. Document any missing values in the data

```{r}
colSums(is.na(rollercoasters))
```

## Categorical variables

6. Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables created using `table()`

- Interpret a number from each resulting table