---
title: "ST558 Homework 5"
author: Ryan Mersereau
format: html
editor: visual
---

# Task 1: Querying an API and a Tidy-Style Function

For this task, we'll need to connect to the news API here: <https://newsapi.org>

1. Use `GET()` from the `httr` package to return information about a topic that has been in the news 
lately

Here I chose to search for articles from last Monday regarding Panthers quarterback Bryce Young
```{r}
#| message: FALSE
#| warning: FALSE
library(httr)
library(jsonlite)
library(tidyverse)
```

```{r}
# Store personal news api key
my_key <- "48e8f55cb8e74372b534d5595a576d81"

# Building the url
id_info <- GET(
  url = "https://newsapi.org/v2/everything",
  query = list(
    q = "bryce-young",
    from = "2026-09-14",
    language = "en",
    pageSize = 50,
    apiKey = my_key
  )
)

str(id_info, max.level = 1)
```

2. Parse what is returned and find your way to the data frame that has the actual article information in it (content). The first column should be a list column!

```{r}
parsed <- fromJSON(rawToChar(id_info$content))

str(parsed, max.level = 1)
```
This returns a list with 3 elements. To specifically look at the articles published,
we can store this as a tibble below.

```{r}
parsed_articles <- as_tibble(parsed$articles)
parsed_articles
```

3. Now write a quick function that allows the user to easily query this API. The inputs
to the function should be the title/subject to search for, a time period to search from, and an
API key.

```{r}
query_function <- function(subject, date, api_key) {
  response <- GET(
    url = "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = date,
      language = "en",
      pageSize = 50,
      apiKey = api_key
    )
  )
  
  parsed <- fromJSON(rawToChar(response$content))
  as_tibble(parsed$articles)
}
```

Lets call this a couple times to show it works

```{r}
bojangles <- query_function("Bojangles", "2026-09-10", my_key)
bojangles
```

```{r}
israel <- query_function("Israel", "2026-08-31", my_key)
israel
```

# Task 2: Summarizing Roller Coaster Data

For this part, we'll use roller coaster data found here: <https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv>

Let's read in this data and convert variables to factors as needed

```{r}
#| message: FALSE
#| warning: FALSE
library(readr)
library(psych)
```


```{r}
rollercoasters <- read_csv("../HW5/635USrollercoasters.csv")
```


```{r}
rollercoasters |>
  select(-c(Boundary, Latitude, Longitude)) |>
  mutate(AgeGroup = factor(AgeGroup, 
                           levels = c("1:Older", "2:Recent", "3:Newest"),
                           labels = c("Older", "Recent", "Newest")),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design)) |>
  rename(CensusRegion = `Census Region`,
         YearOpened = `Year Opened`,
         Inversion_YN = `Inversions?`,
         Inversion_Num = `# of Inversions`) 
```

## Basic EDA

4. Look at how the data is stored and see if everything makes sense.

```{r}
str(rollercoasters)
```

```{r}
psych::describe(rollercoasters)
```


5. Document any missing values in the data

```{r}
colSums(is.na(rollercoasters))
```

## Categorical variables

6. Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables created using `table()`

- Interpret a number from each resulting table

- One-way
```{r}
table(rollercoasters$Type)
```
There are 530 Steel roller coasters in this dataset

- Two-way

```{r}
table(rollercoasters$Type, rollercoasters$Scale)
```

There are 53 Extreme-scale wooden roller coasters in this data

- Three-way
```{r}
table(rollercoasters$Type, rollercoasters$Scale, rollercoasters$AgeGroup)
```

There are 21 Extreme-scale wooden roller coasters in the 'newest' age group.

7. Create a conditional two-way table using `table()`, conditioning on one variables setting

First, we'll do this by subsetting the data (with `filter()`) and then creating the two-way table 

```{r}
filtered_coasters <- rollercoasters |>
  filter(AgeGroup == "3:Newest")

filtered_coasters
```

```{r}
table(filtered_coasters$Type, filtered_coasters$Scale)
```

Here we get the same result as the bottom of the three-way table. Since the data is
filtered for only new coasters, this is a two-way table on Type and Scale filtered on the coaster being new.

We can create a conditional two-way table another way by directly subsetting a three-way table

```{r}
table(rollercoasters$Type, rollercoasters$Scale, rollercoasters$AgeGroup)[, ,3]
```

This is another way to get the same result.

8. Create a two-way contingency table using `group_by` and `summarize()` from `dplyr`. Then use `pivot_wider()` to make the result look more like the output from `table()`

```{r}
rollercoasters |>
  drop_na(Type, Scale) |>
  group_by(Type, Scale) |>
  summarize(count = n()) |>
  pivot_wider(names_from = Type, values_from = Scale)
```

Still doesn't quite look like the results from `table()`, but the values are consistent with our two-way table in **7.**

9. Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

```{r}
#|message: FALSE
#|warning: FALSE

library(ggplot2)
```

**Stacked**

```{r}
g1 <- ggplot(data = rollercoasters |>
               drop_na(Type, Scale), aes(x = Scale, fill = Type))
g1 + geom_bar() + labs(x = "Type", y = "Count", title = "Rollercoaster Types by Scale") 
```

**Side-by-side**

```{r}
ggplot(data = rollercoasters |> drop_na(Type, Scale), aes(x = Scale, fill = Type)) +
  geom_bar(position = "dodge") + 
  labs(x = "Scale", title = "Roller Coaster Types by Scale", y = "Count") +
  scale_fill_discrete("Type")
```

This is better (although the scales should be ordered) From most to least intense it goes:
Extreme, Thrill, Family, Kiddie