---
title: "ST558_HW5_Nowak"
author: "Kayla Nowak"
format: html
editor: visual
---

```{r}
#| message: False
#| warning: False

library(tidyverse)
library(httr)
library(jsonlite)
```

## Task 1: Querying an API in a Tidy-Style Fashion

Connect to the news API here: [newsapi.org](https://newsapi.org/)

### 1.

Use `GET()` from the `httr` package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object).

```{r}
news_pull <- GET("https://newsapi.org/v2/everything?q=Iran&from=2026-09-18&language=en&pageSize=50&apiKey=71ce2ed6f6844a9b8c771a2b3993de20")
```

### 2.

Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
#Check top-level contents
str(news_pull, max.level=1)

#Parse to contents
news_parse <- fromJSON(rawToChar(news_pull$content))
str(news_parse, max.level=1)

#Select articles
news_data <- as_tibble(news_parse$articles)
news_data
```

### 3.

Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
pullFromNews <- function(subject, startdt, key) {
  
  URL <- paste0("https://newsapi.org/v2/everything?q=", as.character(subject),
                "&from=", as.character(startdt),
                "&to=", as.character(Sys.Date()),
                "&language=en&pageSize=50&apiKey=", key)
  
  pulled <- GET(URL)
  #Note: missing function inputs result in R error.
    
  if(pulled$status != 200) { return("Something broke, check requirements.") }
  else {
    parsed <- fromJSON(rawToChar(pulled$content))
    tdata  <- as_tibble(parsed$articles)
    return(tdata)
  }
}

```

Test Function:

```{r}
mykey <- "71ce2ed6f6844a9b8c771a2b3993de20"

#Test 1
pullFromNews(subject="Iran", startdt= "2026-09-18", key=mykey)

#Test 2
pullFromNews(subject="Apple", startdt= "2026-09-01", key=mykey)

#Test 3 (expected to fail)
pullFromNews(subject="Apple", startdt= "2026-08-01", key=mykey)    #Date too old
pullFromNews(subject="Apple", startdt= "2026-09-01", key="123ABC") #Bad Key
```

## Task 2: Summarizing Roller Coaster Data

Working with the roller coaster data. Source: <https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv>

Read-in data and convert variables to factors as needed. No need to use the `Boundary` variable or the `Latitude` and `Longitude` variables in the EDA.

```{r}
rollercoasters <- read_csv("635USrollercoasters.csv") 
names(rollercoasters)

#Preliminary Cleaning:

rollercoasters <- rollercoasters |>
  #drop variables we don't need:
  select(-c(Boundary, Latitude, Longitude)) |>
  #enforce valid column names:
  rename(CensusRegion = `Census Region`,
         YearOpened = `Year Opened`,
         Inversion_YN = `Inversions?`,
         Inversion_Num = `# of Inversions` ) 

names(rollercoasters)
```

### 4.

Look at how the data is stored and see if everything makes sense.

```{r}
str(rollercoasters)
```

-   `Coaster`, `Park`, `City`, `State`, and `Make` are fine to store as character
-   `YearOpened`, `TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, and `Inversion_Num` are appropriately stored as numeric
-   All other variables are likely categorical and may be better as factors.

```{r}
vars_cat <- c("CensusRegion", "AgeGroup", "Scale", "Type", "Design", "Inversion_YN")

for(i in 1:length(vars_cat)){
  rollercoasters |> 
    group_by(across(all_of(vars_cat[i]))) |> 
    summarize(count=n()) |> print()
}
```

Yes, convert all the suspected categorical variables to factors. `AgeGroup` was the only variable that benefited from a labeling change.

```{r}
rollercoasters <- rollercoasters |>
  mutate(CensusRegion = as.factor(CensusRegion),
         AgeGroup = factor(AgeGroup, 
                            levels=c("1:Older", "2:Recent", "3:Newest"),
                            labels=c("Older", "Recent", "Newest")),
         Scale = as.factor(Scale),
         Type= as.factor(Type),
         Design=as.factor(Design),
         InversionYN=as.factor(Inversion_YN))

str(rollercoasters)
```

### 5.

Document any missing values in the data.

```{r}

#Function to count NA within column
sum_na <- function(column){	
  sum(is.na(column))
}

#Apply function, print variables with some missingness
has_na <- rollercoasters |>
  summarize(across(everything(), sum_na)) |>
  pivot_longer(cols=everything(),
               names_to = "Variable",
               values_to = "NA_Count") |>
  filter(NA_Count > 0)

has_na
```


### 6.

Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this. Interpret a number from each table.

**One-Way Table:**

```{r}
table(rollercoasters$Scale)
```

This dataset contains 178 roller coasters at the "Family" scale.

**Two-Way Table:**

```{r}
table(rollercoasters$Scale, rollercoasters$Design)
```

This dataset contains 6 roller coasters with an Inverted design that were classified as the "Family" scale.

**Three-Way Table:**

```{r}
table(rollercoasters$Scale, rollercoasters$Design, rollercoasters$Type)
```

All 6 of the roller coasters with an Inverted design that were classified as the "Family" scale were made from Steel.

### 7.

Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods:

**By subsetting first:**

```{r}
rollercoasters |> filter(Type=="Steel") |> select(Scale, Design) |> table()
#select after filter to accomplish everything in one chain
```

**By subsetting a 3-way:**

```{r}
table(rollercoasters[, c("Scale", "Design", "Type")])[,,1]
#earlier three-way crosstab confirmed 'Steel' was 1st level of Type
```

### 8. 

Create a two-way contingency table using `group_by()` and `summarize()` from dplyr. Then use `pivot_wider()` to make the result look more like the output from `table()`.

```{r}
rollercoasters |>
  group_by(Scale, Design) |>
  summarize(count = n()) |>
  pivot_wider(names_from = Design, 
              values_from = count, 
              values_fill=0)  #fill in zero count for unrepresented pairings
```

*Note: Columns are arranged in a different order than what was produced by the `table()` function. Also, some of the levels from Design were not valid column names. Something to keep in mind.*
