---
title: "ST558_HW5_Yefrid"
author: "Yefrid Cordoba"
format: html
---

## Task 1: Querying an API and a Tidy-Style Function

### 1.

We need to use the `GET()` function from the topic trade

```{r}
#| message: False
#| warning: False

library(httr)
library(jsonlite)
library(tidyverse)
library(readr)
library(gt)
```

Now we obtain the info from the API.

```{r}

info_news <- GET("https://newsapi.org/v2/everything?q=trade&from=2026-08-21&language=en&pageSize=50&apiKey=b83cc47afb8e4f47a5ff9dafee63d05c")
```

We explore the info to determine the actual data that we need to extract.

```{r}
str(info_news)
```

### 2.

We see that the data is under the name `content`.   
We use the `jsonlite` package to get a tibble.

```{r}
parsed <- fromJSON(rawToChar(info_news$content))
team_info <- as_tibble(parsed$articles)
team_info
```

### 3.

Now we will create a funstion that will create the URL to get data from the `News` API.

```{r}
get_news <- function(title, period, key){
  url <- paste(paste0("https://newsapi.org/v2/everything?q=", title),
               paste0("from=", as.character(period)),
               paste0("apiKey=", key),
               sep = "&"
               )
  
  
  info_news <- GET(url = url)
  
  parsed <- fromJSON(rawToChar(info_news$content))
  
  team_info <- as_tibble(parsed$articles)
  
  team_info
  
}
```

Now we can test the function with two different topics and dates

```{r}
get_news("macklemore", "2026-09-15", "b83cc47afb8e4f47a5ff9dafee63d05c")
```

```{r}
get_news("gas", "2026-09-10", "b83cc47afb8e4f47a5ff9dafee63d05c")
```

## Task 2: Summarizing Roller Coaster Data

First we are going to read the data from the roller coaster, and convert the required variables to factor.

```{r}
rc_data <- read_csv("635USrollercoasters.csv") |>
  mutate(State = as.factor(State),
         `Census Region` = as.factor(`Census Region`),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         `Inversions?` = as.factor(`Inversions?`),
         Latitude = NULL,
         Longitude = NULL,
         Boundary = NULL)

```


```{r}
rc_data
```

### 4.

First we need to determine how the data is stored, for that we are going to use `str()` function

```{r}
str(rc_data)
```

As previously transformed, the data that required to be factor were transformed, and the numeric variables are correctly stores as numeric variables

### 5.

And we determine the number of missing values for each of the variables

```{r}
colSums(is.na(rc_data))
```

We can observe that there are missing values for `Drop`, `TopSpeed`, `Duration`, and `Length`, listed in descending order of their missig value counts.

### 6.

- We need to create first a one-way contingency table using the `Type` variable.

```{r}
table(rc_data$Type)
```

Base on the census data, there are 530 roller coasters made of steel.

- Now a two-way contingency table.

```{r}
table(rc_data$Type, rc_data$Scale)
```

Out of the 530 steel roller coasters, 216 are categorized as extreme.

- Now a three-way contingency table

```{r}
table(rc_data$Type, rc_data$Scale, rc_data$AgeGroup)
```

From the older roller coasters, 9 are made of wood and categorized as extreme.

### 7.

First a conditional two-way contingency table, using a subset of the data, and then creating the two-way table
```{r}
sub_rc <- filter(rc_data, `Census Region` == "South")
table(sub_rc$Type, sub_rc$Scale)
```

And then, creating first a three way contingency table, and sub setting it.

```{r}
table(rc_data$Type, rc_data$Scale, rc_data$`Census Region`)[ , , "South"]
```

### 8.

In this step we need to create a two-way contingency table using tidyverse.

```{r}
rc_data |>
  group_by(`Inversions?`, Type) |>
  summarise(count = n(), .groups = "drop") |>
  pivot_wider(names_from = `Inversions?`, values_from = count) 
```

### 9.

Now we need to create a stacked bar plot.

```{r}
ggplot(data = rc_data |>
         group_by(`Inversions?`, Type) |>
         summarise(count = n()),
    aes(x = Type, 
        y = count,
        fill = `Inversions?`)) +
  geom_bar(stat = "identity") +
  labs(title = "Roller Coaster Materials and Presence of Inversions",
       x = "Material") +
  scale_fill_discrete("Inversions")
```

And side by side bar

```{r}
ggplot(data = rc_data,
    aes(x = Type,
        fill = `Inversions?`)) +
  geom_bar(position = "dodge") +
  labs(title = "Roller Coaster Materials and Presence of Inversions",
       x = "Material") +
  scale_fill_discrete("Inversions")
```

