---
title: "ST558_HW5_Nowak"
author: "Kayla Nowak"
format: html
editor: visual
---

```{r}
#| message: False
#| warning: False

library(tidyverse)
library(httr)
library(jsonlite)
```

## Task 1: Querying an API in a Tidy-Style Fashion

Connect to the news API here: [newsapi.org](https://newsapi.org/)

### 1.

Use `GET()` from the `httr` package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object).

```{r}
news_pull <- GET("https://newsapi.org/v2/everything?q=Iran&from=2026-09-18&language=en&pageSize=50&apiKey=71ce2ed6f6844a9b8c771a2b3993de20")
```

### 2.

Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
#Check top-level contents
str(news_pull, max.level=1)

#Parse to contents
news_parse <- fromJSON(rawToChar(news_pull$content))
str(news_parse, max.level=1)

#Select articles
news_data <- as_tibble(news_parse$articles)
news_data
```

### 3.

Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
pullFromNews <- function(subject, startdt, key) {
  
  URL <- paste0("https://newsapi.org/v2/everything?q=", as.character(subject),
                "&from=", as.character(startdt),
                "&to=", as.character(Sys.Date()),
                "&language=en&pageSize=50&apiKey=", key)
  
  pulled <- GET(URL)
  #Note: missing function inputs result in R error.
    
  if(pulled$status != 200) { return("Something broke, check requirements.") }
  else {
    parsed <- fromJSON(rawToChar(pulled$content))
    tdata  <- as_tibble(parsed$articles)
    return(tdata)
  }
}

```

Test Function:

```{r}
mykey <- "71ce2ed6f6844a9b8c771a2b3993de20"

#Test 1
pullFromNews(subject="Iran", startdt= "2026-09-18", key=mykey)

#Test 2
pullFromNews(subject="Apple", startdt= "2026-09-01", key=mykey)

#Test 3 (expected to fail)
pullFromNews(subject="Apple", startdt= "2026-08-01", key=mykey)    #Date too old
pullFromNews(subject="Apple", startdt= "2026-09-01", key="123ABC") #Bad Key
```
